---
title: "Generate a null distribution for rail and roads"
author: "Nadege Belouard"
date: "1/18/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(spData)
library(RVAideMemoire)
library(dplyr)
library(gridExtra)
library(magrittr)
library(purrr)
library(here)
library(FSA)

```


# 1. Generate random dispersal distribution

Load files
```{r read files}

# Load points from chull
grid_data <- read.csv(file.path(here(), "exported-data", "grid_chull_unique.csv"))
head(grid_data)
dim(grid_data)

#Create a sf layer of positive points
slfEstab <- grid_data %>% filter(slf_established == TRUE)
slfEstab_layer <- st_as_sf(x = slfEstab, coords = c("longitude_rounded", "latitude_rounded"), crs = "EPSG:4269", remove = F) %>%
  st_transform(slfEstab_layer, crs = "ESRI:102010")

# Create a US map 
US <- st_read("C:/Users/labuser/Documents/Postdoc_SLF/SLF_Dispersal/data/raw_data/states/gadm36_Cont_USA_county/gadm36_Cont_USA_county.shp", crs = "EPSG:4326", quiet = T)
st_crs(US)
US <- st_transform(US, crs = "ESRI:102010")


# Visualize the SLF data
ggplot(data = US, fill = "white") +
  geom_sf() +
  geom_sf(data = slfEstab_layer, col = "blue") +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)
```


Generate all possible random coordinates
```{r generate coordinates}

# (1) Calculate all potential coordinates
maxlat <- max(slfEstab$latitude_rounded)
minlat <- min(slfEstab$latitude_rounded)
maxlong <- max(slfEstab$longitude_rounded)  
minlong <- min(slfEstab$longitude_rounded)

Seqlat <- seq(from = minlat, to = maxlat, by = 1/111)
Seqlong <- seq(from = minlong, to = maxlong, by = 1/85)
Coordinates <- expand.grid(latitude = Seqlat, longitude = Seqlong)
Coordinates <- st_as_sf(x = Coordinates, coords = c("longitude", "latitude"), crs = "EPSG:4269", remove = F) %>%
  st_transform(Coordinates, crs = "ESRI:102010")


# Map these points
ggplot(data = US, fill = "white") +
  geom_sf() +
  geom_sf(data = Coordinates, alpha = 0.5) +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)


# (2) Keep only the coordinates that are in the minimum convex polygon
chull <- st_read(file.path(here(), "figures", "GIS", "chull.shp"), quiet = T)
Coordinates_chull <- st_intersection(Coordinates, hull)

ggplot(data = US, fill = "white") +
  geom_sf() +
  geom_sf(data = Coordinates_chull, alpha = 0.5) +
  geom_sf(data = slfEstab_layer, col = "blue") +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)


# (3) Keep only those that are on land!
Coordinates_land <- st_intersection(Coordinates_chull, US)

random_coord <- ggplot(data = states, fill = "white") +
  geom_sf() +
  geom_sf(data = Coordinates_land, alpha = 0.5) +
  geom_sf(data = chull, alpha = 0.5, fill = "white") + 
  # geom_point(data = slfEstab_layer,
             # aes(x = longitude_rounded, y = latitude_rounded), col = "black") +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(-82, -72), ylim = c(37, 43)) +
  theme_classic()


# Save the file
Coordinates_land %<>% select(latitude, longitude, geometry)
st_write(Coordinates_land, file.path(here(), "exported-data", "slf_sim_coordinates.shp"), driver = "ESRI Shapefile")

```


# 2. Calculate distances for the grid of random points 
(step done on the HPC - computationally intensive). Can reduce the time by removing points for which we already have data from the previous vignette (observed data)
```{r calculate distances}

Coordinates_land <- st_read(file.path(here(), "exported-data", "slf_sim_coordinates.shp"), quiet = T)

missing_distances <- setdiff(Coordinates_land, grid_chull_unique)

# Create rows for distances
missing_distances %<>% add_column(DistToRail = NA,
                          DistToRoad = NA)

#Calculate their distance to transport infrastructures
for (j in 1:length(Coordinates_land$DistToRail)){ 
  # Print the row being considered
  print(j)
  
  point_clip <- st_buffer(Coordinates_land[j,], dist = 50000)
  
  # Calculate distance to the closest railroad
  rail_clip <- st_intersection(rail, point_clip)
  dist_rail <- st_distance(x = Coordinates_land[j,], y = rail_clip)
  Coordinates_land$DistToRail[j] <- min(dist_rail)
  
  # Calculate distance to the closest major road
  road_clip <- st_intersection(road, point_clip)
  dist_road <- st_distance(x = Coordinates_land[j,], y = road_clip)
  Coordinates_land$DistToRoad[j] <- min(dist_road)
  
  rm(point_clip, rail_clip, road_clip)
  
  # Stores the result every 1,000 iteration to avoid loosing all the data in case of crash
  if (i %% 1000 == 0){
    st_write(Coordinates_land, file.path(here(), "exported-data", "slf_sim_dist_transports.shp"), row.names = F)
  }
}


# Save file
st_geometry(Coordinates_land) <- NULL
write.csv(Coordinates_land, file.path(here(), "exported-data", "slf_sim_dist_transports.csv"), row.names = F)
```


# 3. Create random datasets
Extract 9,999 random datasets of the same number of points as the observed data

Full dataset
```{r select random full datasets}

# Load distances and points
Sim_dist_transport <- read.csv(file.path(here(), "exported-data", "slf_sim_dist_transport.csv"))
head(Coordinates_land)

slf_uptodate <- read.csv(file.path(here(), "exported-data", "slf_uptodate.csv"), h=T)

# Size of the dataset to be sampled
jumpers_full = dim(slf_uptodate %>% filter(Category_full == "Jumpers"))[1]


for (i in 1:1){
  #Generate a set of coordinates
  Random_coordinates <- Coordinates_land[sample(nrow(Coordinates_land), size = jumpers_full, replace = F),] %>%
    add_column(Category_full = "Jumpers")
  
  #Calculate the mean and median distance per simulation
  Random_means <- Random_coordinates %>% group_by(Category_full) %>% 
    summarise_at(vars(starts_with("DistTo")), list(mean = mean, median = median)) %>% 
    add_column(Simulation = i)
  
  #Save the table with the simulation number
  if (i == 1){
    Simulations <- Random_means
  } else {
    Simulations <- bind_rows(Simulations, Random_means)
  }
  
  if (i %% 100 == 0){ print(i)}
} 

# write.csv(Simulations, "./exported-data/Distances_random distrib/SimulatedMeans_fulluptodate.csv", row.names = F)
write.csv(Simulations, file.path(here(), "exported-data", "slf_sim_full_transports.csv"), row.names = F)
```

Rarefied dataset
```{r sample rarefied dataset}
# Load distances and points
Sim_dist_transport <- read.csv(file.path(here(), "exported-data", "slf_sim_dist_transport.csv"))
head(Coordinates_land)

slf_uptodate <- read.csv(file.path(here(), "exported-data", "slf_uptodate.csv"), h=T)

# Size of the dataset to be sampled
jumpers_rarefied = dim(slf_uptodate %>% filter(Category_rare == "Jumpers"))[1]

for (i in 1:9999){
  #Generate a set of coordinates
  Random_coordinates <- Coordinates_land[sample(nrow(Coordinates_land), size = jumpers_rarefied, replace = F),] %>%
    add_column(Category_rarefied = "Jumpers")

  #Calculate the mean and median distance per simulation
  Random_means <- Random_coordinates %>% group_by(Category_rarefied) %>% 
    summarise_at(vars(starts_with("DistTo")), list(mean = mean, median = median)) %>% 
    add_column(Simulation = i)
  
  #Save the table with the simulation number
  if (i == 1){
    Simulations <- Random_means
  } else {
    Simulations <- bind_rows(Simulations, Random_means)
  }
  
  if (i %% 100 == 0){ print(i)}
} 

write.csv(Simulations, file.path(here(), "exported-data", "slf_sim_rarefied_transports.csv"), row.names = F)

```


Map points
```{r map of random points for jumpers, fig.width= 6, fig.height=7, eval = FALSE}

# Example of random coordinates
ggplot(data = states) +
  geom_sf(fill = "white") + 
  geom_point(data = Random_jumpers_full, aes(x = longitude, y = latitude), col = "blue") +
  labs(x = "Longitude", y = "Latitude")+
  geom_text(data = states, aes(X, Y, label = code), size = 5) +
  theme(legend.position="top") +
  coord_sf(xlim = c(-82, -72), ylim = c(37, 43), expand = FALSE)
```



# 4. Visualize results

Load observed data
```{r load observed data}

# Convert the observed data to long format, get the average values
slf_uptodate <- read.csv(file.path(here(), "exported-data", "slf_uptodate.csv"))
dim(slf_uptodate) #32,743 rows

# Convert to long format
slf_uptodate_long <- slf_uptodate %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")

# Calculate mean and median
slf_obsmeans <- slf_uptodate_long %>% 
  group_by(Category_full, DistanceType) %>% 
  summarise(MeanDistance = mean(DistanceValue),
            MedianDistance = median(DistanceValue))
```

## Full dataset
Calculate mean and median for simulated data
```{r check results}

# Load the simulated values
Simulations_full <- read.csv(file.path(here(), "exported-data", "slf_sim_full_transports.csv"))

# Convert simulations to long format
Simulations_full_long <- Simulations_full %>% pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")

# Select only the means
Simulations_full_long_mean <- Simulations_full_long %>%
  filter(DistanceType %in% c("DistToRail_mean", "DistToRoad_mean"))
Simulations_full_long_mean$DistanceType <- as.factor(Simulations_full_long_mean$DistanceType)
levels(Simulations_full_long_mean$DistanceType) <- c("DistToRail", "DistToRoad")

# Select only the medians
Simulations_full_long_median <- Simulations_full_long %>% filter(DistanceType %in% c("DistToRail_median", "DistToRoad_median"))
Simulations_full_long_median$DistanceType <- as.factor(Simulations_full_long_median$DistanceType)
levels(Simulations_full_long_median$DistanceType) <- c("DistToRail", "DistToRoad")
```

Visualize results
```{r mean and median for observed values}

#Plot means
random_transport <- ggplot() +
  geom_histogram(data = Simulations_full_long_mean %>% filter(Category_full == "Jumpers"), 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_full), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans %>% filter(Category_full == "Jumpers"),
             mapping = aes(xintercept = MeanDistance/1000, col = Category_full), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

ggsave(file.path(here(), "figures", "jump_transports", "bootstrap_transports_fullmeans.jpg"), random_transport, width = 10, height = 3)



# Plot medians
random_transport <- ggplot() +
  geom_histogram(data = Simulations_full_long_median %>% filter(Category_full == "Jumpers"), 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_full), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans %>% filter(Category_full == "Jumpers"),
             mapping = aes(xintercept = MedianDistance/1000, col = Category_full), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

random_transport

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_fullmedianschull_jumpers.jpg"), random_transport, width = 10, height = 3)
``` 

Count simulations
```{r count simulations}
# COUNT HOW MANY SIMULATIONS ARE SMALLER THAN THE OBS VALUE
obsrail <- slf_obsmeans %>% filter(Category_full == "Jumpers", DistanceType == "DistToRail") %>% pull(MedianDistance) 
dim(Simulations_full_long_median %>% filter(Category_full == "Jumpers", DistanceType == "DistToRail", DistanceValue < obsrail))[1]
#0

obsroad <- slf_obsmeans %>% filter(Category_full == "Jumpers", DistanceType == "DistToRoad") %>% pull(MedianDistance) 
dim(Simulations_full_long_median %>% filter(Category_full == "Jumpers", DistanceType == "DistToRoad", DistanceValue < obsrail))[1]
#0
```


## Rarefied dataset
Calculate mean and median for simulated data
```{r count simulations rarefied}

# Load the simulated values
Simulations_rare <- read.csv(file.path(here(), "exported-data", "slf_sim_rarefied_transports.csv"))

# Convert simulations to long format
Simulations_rare_long <- Simulations_rare %>% pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")
Simulations_rare_long %<>% rename(Category_rare = Category_rarefied) 
Simulations_rare_long$Category_rare <- factor(Simulations_rare_long$Category_rare, levels = c("Jumpers", "Diffusers", "Negatives"))


# Select only the means
Simulations_rare_long_mean <- Simulations_rare_long %>% filter(DistanceType %in% c("DistToRail_mean", "DistToRoad_mean", "DistToAirport_mean", "DistToPort_mean"))
Simulations_rare_long_mean$DistanceType <- as.factor(Simulations_rare_long_mean$DistanceType)
levels(Simulations_rare_long_mean$DistanceType)
levels(Simulations_rare_long_mean$DistanceType) <- c("DistToAirport", "DistToPort", "DistToRail", "DistToRoad")

# Select only the medians
Simulations_rare_long_median <- Simulations_rare_long %>% filter(DistanceType %in% c("DistToRail_median", "DistToRoad_median", "DistToAirport_median", "DistToPort_median"))
Simulations_rare_long_median$DistanceType <- as.factor(Simulations_rare_long_median$DistanceType)
levels(Simulations_rare_long_median$DistanceType)
levels(Simulations_rare_long_median$DistanceType) <- c("DistToAirport", "DistToPort", "DistToRail", "DistToRoad")
```

Visualize results
```{r visualize rare}

#Plot means
random_transport <- ggplot() +
  geom_histogram(data = Simulations_rare_long_mean %>% filter(Category_rare == "Jumpers"), 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_rare), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans,
             mapping = aes(xintercept = MeanDistance/1000, col = Category_rare), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~Category_rare + DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_raremeanschull.jpg"), random_transport, width = 10, height = 9)


# Plot median
random_transport <- ggplot() +
  geom_histogram(data = Simulations_rare_long_median %>% filter(Category_rare == "Jumpers"), 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_rare), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans %>% filter(Category_rare == "Jumpers"),
             mapping = aes(xintercept = MedianDistance/1000, col = Category_rare), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_raremedianschull_jumpers.jpg"), random_transport, width = 10, height = 3)
```

Count simulations
```{r count simulations rare}

# COUNT HOW MANY SIMULATIONS ARE SMALLER THAN THE OBS VALUE
obsrail <- slf_obsmeans %>% filter(Category_rare == "Jumpers", DistanceType == "DistToRail") %>% pull(MedianDistance) 
dim(Simulations_rare_long_median %>% filter(Category_rare == "Jumpers", DistanceType == "DistToRail", DistanceValue < obsrail))[1]
#0

obsroad <- slf_obsmeans %>% filter(Category_rare == "Jumpers", DistanceType == "DistToRoad") %>% pull(MedianDistance) 
dim(Simulations_rare_long_median %>% filter(Category_rare == "Jumpers", DistanceType == "DistToRoad", DistanceValue < obsrail))[1]
#148
``` 


## Check distribution of observed distances
```{r check distribution of observed distances}
slf_uptodate <- read.csv(file.path(here(), "exported-data", "slf_uptodate.csv"))
dim(slf_uptodate) #43,801 rows

slf_uptodate_long <- slf_uptodate %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")

slf_uptodate_long$Category_full <- factor(slf_uptodate_long$Category_full, levels = c("Jumpers", "Diffusers", "Negatives"))

distrib <- ggplot(slf_uptodate_long, aes(x =  DistanceValue/1000, y = ..density.., fill = Category_full), binwidth = 0.1) +
  geom_histogram() +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~Category_full+DistanceType, scales = "free") +
  theme_classic() 


ggsave(file.path(here(), "figures", "vignette_transports", "distribution_observed_chull.jpg"), distrib, width = 6, height = 6)
```


# 5. Calculate effect sizes

## Full dataset
Load observed data
```{r load observed data}
# Observed data
slf_uptodate <- read.csv(file.path(here(), "exported-data", "points_chull.csv"))
slf_cat_long <- slf_uptodate %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")

slf_obsmeans <- slf_cat_long %>% 
  group_by(Category_full, DistanceType) %>% 
  summarise(MeanDistance = mean(DistanceValue)#,
            # MedianDistance = median(DistanceValue)
            )
```

Load simulated data
```{r effect sizes}
# Simulated means
MeanDistances <- read.csv(file.path(here(), "exported-data",  "SimulatedMeans_fullchull.csv"))
names(MeanDistances)
head(MeanDistances)

#Modify dataset to get distribution of distances with a column for the type of distance, and one column for the type of dataset (Full or Rarefied)
random_long <- MeanDistances %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")
```

Means
```{r long mean}

random_long_mean <- random_long %>% filter(grepl("_mean", DistanceType))
head(random_long_mean)
random_long_mean$DistanceType <- as.factor(random_long_mean$DistanceType)
levels(random_long_mean$DistanceType)
levels(random_long_mean$DistanceType) <- gsub(pattern = "_mean", replacement = "", x = levels(random_long_mean$DistanceType))


head(random_long_mean)
summary_sim_full <- random_long_mean %>% group_by(Category_full, DistanceType) %>% summarise(mean_sim = mean(DistanceValue), sd_sim = sd(DistanceValue))
head(summary_sim_full)


summary_full <- merge(summary_sim_full, slf_obsmeans)
summary_full %<>% rename(mean_obs = MeanDistance) %>% 
  mutate(effect_size = (mean_obs - mean_sim)/sd_sim)
write.csv(summary_full, "summary_simulations_full_transports.csv")


Category = as.vector(unique(summary_full$DistanceType))
str(Category)
Properties = c("Primary airports (more than 10,000 passenger boardings each year, FAA, USGS National Transportation Dataset)",
               "Ports (USGS Science-Base Catalog",
               "Railroads (USGS National Transportation Dataset)",
               "Major roads (primary and secondary roads as defined by the MAF/TIGER Feature Classification, US Census Bureau)")
Names <- data.frame(DistanceType = Category, Properties = Properties)


summary_full_cat <- merge(summary_full, Names)
write.csv(summary_full_cat, "summary_simulations_full_transports.csv")


# GRAPH
ggplot(summary_full_cat, aes(y = effect_size, x = DistanceType)) + 
  geom_point(aes(col = Category_full)) +
  # scale_fill_brewer(palette = "Dark2") +
  # facet_wrap(~Category_full, scales = "free_y", ncol = 1) +
             # labeller = labeller(DistanceType = 
  #   c("DistToAirport" = "Airport",
  #     "DistToBoats" = "Boating",
  #     "DistToBottling" = "Bottling plants",
  #     "DistToColleges" = "Colleges",
  #     "DistToDistrib" = "Distribution centers",
  #     "DistToGarages" = "Garages",
  #     "DistToIntermodal" = "Intermodal platforms",
  #     "DistToMail" = "Mail carriers",
  #     "DistToMarket" = "Farmers market",
  #     "DistToMoving" = "Moving companies",
  #     "DistToPeople" = "Popular destinations",
  #     "DistToPort" = "Port",
  #     "DistToRail" =  "Railroad",
  #     "DistToRoad" = "Major road",
  #     "DistToTruckStop" = "Truck stops",
  #     "DistToWineries" = "Wineries",
  #     "DistToWood" = "Wood-related activities"))) +
  # # coord_cartesian(ylim=c(0, 30000)) +
  # xlab("Category") + ylab("Distance to the nearest... (km)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Medians
```{r median}

random_long_median <- random_long %>% filter(grepl("_median", DistanceType))
head(random_long_median)
random_long_median$DistanceType <- as.factor(random_long_median$DistanceType)
levels(random_long_median$DistanceType)
levels(random_long_median$DistanceType) <- gsub(pattern = "_median", replacement = "", x = levels(random_long_median$DistanceType))


head(random_long_median)
summary_sim_medianfull <- random_long_median %>% group_by(Category_full, DistanceType) %>% summarise(median_sim = median(DistanceValue), sd_sim = sd(DistanceValue))
head(summary_sim_medianfull)

summary_medianfull <- merge(summary_sim_medianfull, slf_obsmedians)
summary_medianfull %<>% rename(median_obs = MedianDistance) %>% 
  mutate(effect_size = (median_obs - median_sim)/sd_sim)
write.csv(summary_medianfull, "summary_simulations_median_full_transports.csv")


summary_medianfull_cat <- merge(summary_medianfull, Names)
write.csv(summary_medianfull_cat, "summary_simulations_median_full_transports.csv")
```


## Rarefied dataset

Load observed data
```{r load observed data}
# Observed data
slf_uptodate <- read.csv(file.path(here(), "exported-data", "slf_uptodate.csv"))
slf_cat_long <- slf_uptodate %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")

slf_obsmeans <- slf_cat_long %>% 
  group_by(Category_rare, DistanceType) %>% 
  summarise(MeanDistance = mean(DistanceValue)#,
            # MedianDistance = median(DistanceValue)
            )
```

Load simulated data
```{r effect sizes rarefied}

# Simulated means
MeanDistances <- read.csv(file.path(here(), "exported-data",  "SimulatedMeans_rarefiedchull.csv"))
names(MeanDistances)
head(MeanDistances)

#Modify dataset to get distribution of distances with a column for the type of distance, and one column for the type of dataset (Full or Rarefied)
random_long <- MeanDistances %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")
```

Means 
```{r means rare}
random_long_mean <- random_long %>% filter(grepl("_mean", DistanceType))
head(random_long_mean)
random_long_mean$DistanceType <- as.factor(random_long_mean$DistanceType)
levels(random_long_mean$DistanceType)
levels(random_long_mean$DistanceType) <- gsub(pattern = "_mean", replacement = "", x = levels(random_long_mean$DistanceType))


head(random_long_mean)
summary_sim_rare <- random_long_mean %>% group_by(Category_rarefied, DistanceType) %>% summarise(mean_sim = mean(DistanceValue), sd_sim = sd(DistanceValue))
head(summary_sim_rare)


summary_rare <- merge(summary_sim_rare, slf_obsmeans)
summary_rare %<>% rename(mean_obs = MeanDistance) %>% 
  mutate(effect_size = (mean_obs - mean_sim)/sd_sim)
write.csv(summary_rare, "summary_simulations_rare.csv")


summary_rare_cat <- merge(summary_rare, Names)
write.csv(summary_rare_cat, "summary_simulations_mean_rare_transports.csv")
```

Medians 
```{r medians rare}

random_long_median <- random_long %>% filter(grepl("_median", DistanceType))
head(random_long_median)
random_long_median$DistanceType <- as.factor(random_long_median$DistanceType)
levels(random_long_median$DistanceType)
levels(random_long_median$DistanceType) <- gsub(pattern = "_median", replacement = "", x = levels(random_long_median$DistanceType))


head(random_long_median)
summary_sim_medianrare <- random_long_median %>% group_by(Category_rarefied, DistanceType) %>% summarise(median_sim = median(DistanceValue), sd_sim = sd(DistanceValue))
head(summary_sim_medianrare)


summary_medianrare <- merge(summary_sim_medianrare, slf_obsmedians)
summary_medianrare %<>% rename(median_obs = MedianDistance) %>% 
  mutate(effect_size = (median_obs - median_sim)/sd_sim)
write.csv(summary_medianrare, "summary_simulations_median_rare.csv")

summary_rare_mediancat <- merge(summary_medianrare, Names)
write.csv(summary_rare_mediancat, "summary_simulations_median_rare_transports.csv")
```


The histogram represents the distribution of distances under the null hypothesis of random dispersal of jumpers, for each type of transport. The black vertical lines indicate the significance limits. An observed value situated outside of these vertical lines leads to the rejection of the null hypothesis. The red line indicates the average distance to transports observed in our dataset.  
For all three types of transports, the observed location of jumpers is significantly closer to transports than random.  


\newpage

# 6. Results and conclusion

In this vignette, we discovered that:  

(1) SLF populations, both from jump events and diffusive spread, are not located randomly, but very significantly close to transport infrastructures: roads, rail and airports. The difference to the random distribution is the highest for railroads, then roads, then airports.  *==> SLF presence is tightly linked to transport infrastructures*

(2) Jump events are situated even closer to transport infrastructures than the other SLF populations, on average 444, 659 and 6,566 m away from major roads, rails and airports, respectively. *==> the establishment of new satellite populations is even more linked to transport infrastructures than diffusive spread*

(3) On the other hand, locations where SLF are not found are situated further than random from transport infrastructures. *==> SLF are unlikely to establish far from transport infrastructures*


Given these results, it is very likely that there is a causal relationship between SLF presence and transport infrastructures. SLF are likely transported by vehicles or their content, either as egg masses laid onto random materials, or at another life stage that crawls or flies on vehicles.
It is likely that some transports are more involved in SLF dispersal than others. Correlation is not causation: the proximity to some transports might appear significantly related to SLF presence only because these transports are found in overall highly connected areas. This might be the case for airports, because the correlation between SLF and airports appears looser than with roads or rails. Airports are typically in areas well connected by roads and rails. More, it is assumed that adult SLF cannot survive a flight.


In the next vignette, we are going to apply this knowledge to the spatial analysis of the northeastern US, by looking at areas considered as high risk of invasion (their proximity to SLF populations and to transport infrastructures), and projecting areas likely to be invaded sooner or later.


