---
title: "old code"
author: "Nadege Belouard"
date: "3/30/2022"
output: html_document
---
# 2. Yearly radius of the invasion (DEPRECATED)


To estimate the spread of the SLF, we extract for each year the radius of the invasion, defined as the maximum distance of a survey with established SLF to the introduction point. Median or mean distances are not informative here because surveys are preferentially conducted towards the invasion front, and thus bias the distribution of distances.

```{r advance of the invasion, fig.cap = "Yearly radius of the invasion", fig.height = 5, fig.width = 9, warning = F, echo = params$display, message = FALSE}

# #Computes the maximum distance to the introduction site for each year, status and quadrant
# advance_tab <- grid_data_DP12_rotate0 %>% filter(Status == "Established") %>%
#   group_by(Status, bio_year, portion_name) %>% 
#   summarise(MaxDist = max(DistToIntro))
# 
# # same thing but in a figure
# advance_graph <- ggplot(data = advance_tab, aes(x = bio_year , y = MaxDist)) +
#   geom_bar(stat="identity", fill = "darkgrey", col = "black") +
#   ylab("Invasion radius (km)")+
#   # ylim(c(0,1000)) +
#   xlab("Year") +
#   facet_wrap(~portion_name, ncol = 4) +
#   theme_classic()
# 
# advance_graph

```

```{r save graph advance radius, eval = params$savefiles, echo = params$display}
# ggsave("../figures/vignette_quadrants/advance_graph.jpg", advance_graph, width = 8, height = 6)
```

The invasion radius increased regularly until 2017 (Figure 2). From 2018, there is a steep increase in the invasion radius, especially in the westernmost disk portions, denoting the apparition of dispersal jumps. Note that in 2020, the invasion radius did not increase dramatically, and even decreased in the NNE and ESE portions. This might be due to fewer surveys being conducted that year because of the covid19 pandemic, or be an actual biological pattern. The accuracy of these values can be investigated by checking whether negative surveys are found further than positive surveys, to make sure that SLF were not missed.


\newpage

## Exploration of histograms of distances to the introduction site (DEPRECATED)

Let's have a look at the distribution of the distances of positive vs negative surveys to the introduction point (Figure 3).

```{r histogram of spread distances per year, fig.cap = "Distribution of the distance between SLF populations and the introduction site, per disk portion and per year", fig.height = 20, fig.width = 20, warning = F, echo = params$display}

# histogram_distances <- ggplot(data = grid_data_rotate0 %>% filter(bio_year %in% c(2015:2020), Status != "Present"), aes(x = DistToIntro)) + 
#   geom_histogram(aes(x = DistToIntro, fill = Status),
#                  breaks = seq(0,500,10)) +
#   facet_wrap(~portion_name + bio_year, ncol = 6) +
#   scale_fill_manual(values = c("grey", "red")) +
#   xlab("Distance from the introduction site (km)") +
#   ylab("Number of survey locations") +
#   theme(plot.title = element_text(hjust = 0.5, size = 12), legend.position = "top")
# 
# histogram_distances
```



```{save map disk portions, eval = params$savefiles, echo = params$display}
# ggsave("../figures/vignette_quadrants/histogram_distances.jpg", histogram_distances, width = 20, height = 20)
```

The fact that surveys with SLF undetected are always recorded further than detection events indicates that we can be be fairly confident that the spread of the SLF is accurately monitored. It is also the case in 2020 in the NNE and ESE portions. We can also see that the distribution of established populations is sometimes discontinuous, with gaps where populations were not detected. Detections that appear after the first gap are likely the result of jump dispersal, i.e. human-vectored transportation of SLF in new locations (secondary introductions). 

We can further understand the yearly spread of the SLF by distinguishing diffusive spread (the continuous progress of the invasion) and jump dispersal (long-distance, human-vectored dispersal). We calculated, for each year, the limit of diffusive spread by finding the first 10-mile gap in the distribution of the distances to the introduction site. 
