---
title: "Figures of jump list for the manuscript"
author: 
- Nadege Belouard^[Temple University, nadege.belouard@temple.edu]
- Sebastiona De Bona^[Temple University, seba.debona@temple.edu]
- Jocelyn E. Behm^[Temple University, jebehm@temple.edu]
- Matthew R. Helmus^[Temple University, mrhelmus@temple.edu]
date: "4/8/2022"
output:
  html_document:
    toc: TRUE
    toc_depth: 3
  pdf_document:
    toc: TRUE
    toc_depth: 2
params:
  display: FALSE
  run: TRUE
  loadfiles: FALSE
  savefiles: TRUE
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup for rendering, include = F, messages = F, warning = F}

# attaching necessary packages
library(tidyverse)
library(magrittr)
library(sf)
library(maps)
library(DescTools)
library(reshape2)
library(ggplot2)
library(here)
library(slfjumps)
library(leaflet)
library(purrr)
library(geosphere)
library(cowplot)

sf::sf_use_s2(FALSE)
```

## States map
```{r states names and centroid for global map, message = FALSE, warning = FALSE, echo = params$display}
# extracts a map of the States and recodes state labels to show the two-letter code rather than the full state name.

# obtaining simple feature objects for states and finding centroids for label positioning
states <- sf::st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))
# sf::st_as_sf(maps::map("county", plot = TRUE, fill = FALSE))
states <- cbind(states, st_coordinates(st_centroid(states)))

# making table key for state 2-letter abbreviations
# the vectors state.abb and state.name contains strings of all
# US states and abbreviations
state_abbr <- tibble(state.name = str_to_lower(state.name), state.abb) %>%
  left_join(tibble(ID = states$ID), ., by = c(ID = "state.name")) %>%
  mutate(state.abb = replace_na(state.abb, ""))

# adding 2-letter codes to sf
states$code <- state_abbr$state.abb

ggplot() +
    geom_sf(data = states, alpha = 0) +
  geom_text(data = states, aes(X, Y, label = code), size = 2) 

```


## Load files
```{r load datasets for analysis}
Jumps <- read.csv(file.path(here(), "exported-data", "jumps_full_rarefied.csv"))
Jumps %<>% mutate(Rarefied = ifelse(is.na(Rarefied), "Full", "Rarefied"))
dim(Jumps)
Jumps$Rarefied <- factor(Jumps$Rarefied, levels = c( "Full", "Rarefied"))

Thresholds <- read.csv(file.path(here(), "exported-data", "thresholds.csv"))

slftiny <- read.csv(file.path(here(), "exported-data", "slftiny.csv"), h=T)
grid_data <- read.csv(file.path(here(), "exported-data", "grid_data.csv"), h=T)

centroid <- data.frame(longitude_rounded = -75.675340, latitude_rounded = 40.415240)

jump_groups <- read.csv(file.path(here(), "exported-data", "jump_groups.csv"))

hull <- st_read(file.path(here(), "figures", "GIS", "chull.shp"), quiet = T)
```



#Figure S1. Map of jumps

All points
```{r map data}
# Big map
surveys <- ggplot(data = states) +
  geom_sf(fill = "white") +
  geom_point(data = slftiny,
             aes(x = longitude, y = latitude), col = "gray", size = 1) +
  geom_point(data = grid_data %>% filter(slf_established == TRUE),
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", size = 1) +
  geom_sf(data = states, fill = "white", alpha = 0) +
  labs(x = "Longitude", y = "Latitude") +
  # coord_sf(xlim = c(-90, -70), ylim = c(33, 45)) +
  theme_classic() +
  theme(legend.position="bottom", legend.key = element_rect(fill = "white", colour = NA)) 

surveys

ggsave(file.path(here(), "figures", "manuscript figures", "S1. points_all.jpg"), surveys, width = 6, height = 6)
```

Zoomed map
```{r zoomed map S1}
head(grid_data)
grid_data %<>% mutate(SLF = ifelse(slf_established == T, "Present", "Absent"))

surveys <- ggplot(data = states) +
  geom_sf(fill = "white") +
  geom_point(data = grid_data,
             aes(x = longitude_rounded, y = latitude_rounded, col = SLF), size = 0.5) +
    geom_point(data = grid_data %>% filter(slf_established == TRUE), aes(x = longitude_rounded, y = latitude_rounded), col = "black", size = 0.5) +
    geom_sf(data = hull, alpha = 0, fill = "white") +
  scale_color_manual(values = c("gray", "black")) +
  geom_sf(data = states, fill = "white", alpha = 0) +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(-83, -72), ylim = c(37, 44)) +
  theme_classic() +
  theme(legend.position="bottom", legend.key = element_rect(fill = "white", colour = NA)) 

surveys

ggsave(file.path(here(), "figures", "manuscript figures", "S1. points.jpg"), surveys, width = 6, height = 6)
```



# Figure S2: optimization

```{r create figure of optimization, eval = F}

sectors = c(4,8,12,16,20,24,28,40,60,80,100)
rotations = c(1,2,3,4,5,10,15,20)
centroid <- c(-75.675340, 40.415240)
optim <- data.frame(sectors = 0,
          rotations = 0,
          jumps = 0)
  
optim_list <- data.frame(s = NULL,
                         r = NULL,
                         DistToIntro = NULL,
                         bio_year = NULL,
                         latitude_rounded = NULL,
                         longitude_rounded = NULL,
                         slf_present = NULL,
                         slf_established = NULL,
                         theta = NULL,
                         DistToSLF = NULL)


dim(grid_data %>% filter(slf_established == TRUE, bio_year %in% c(2014:2020)))


for (s in sectors){
  for (r in rotations){
    i = 1
    print(paste0("Sectors: ", s, ", rotations: ", r))
    slfdata <- attribute_sectors(grid_data, nb_sectors = s, centroid = centroid, rotation = r)
    slfdata_long <- slfdata %>% 
      pivot_longer(cols = starts_with("rotation"), names_to = "rotation_nb", 
                   values_to = "sectors_nb", names_prefix = "rotation", 
                   names_transform = list(rotation_nb = as.integer)) 
    Results <- find_jumps(slfdata_long, gap_size = 15, bio_year = c(2014:2020))
    Results$Jump %<>% add_column(r = r, s = s)
    optim_list <- rbind(optim_list, Results$Jump)
    i = i + 1
  }
}

# List of jumps per set of parameters
write.csv(optim_list, file.path(here(), "tables", "jumps_optimization_list.csv"), row.names = F)


```

Simple plot optimization
```{r simple optimization plot}

optim_sum <- optim_list %>% group_by(s, r) %>% summarise(jumps = n())

# Plot it
optim_plot <- ggplot(data = optim_sum, 
                     aes(x = as.factor(r), y = jumps)) +
  geom_bar(lwd = .05, stat = "identity") +
  scale_fill_manual(values = c("#009E73", "#0072B2", "firebrick3", "gold2", "black")) +
  xlab("Number of rotations") +
  ylab("Number of jumps") +
  facet_wrap(~as.factor(s), nrow = 1) +
  guides(fill = guide_legend("Biological year", reverse = T), alpha = guide_legend("Dataset")) +
  theme_classic() +
  ggtitle("Number of sectors") +
  theme(legend.position = "bottom", text = element_text(size = 12), axis.text.x = element_text(size = 5, angle = 90), plot.title = element_text(size = 12, hjust = 0.5))

optim_plot
```

Add rarefied data on plot
```{r optim plot with rarefied data}

# Run attribute_groups and rarefy on this
rarefied_list <- data.frame(bio_year = NULL,
                          latitude_rounded = NULL,
                          longitude_rounded = NULL,
                          s = NULL,
                          r = NULL,
                          Rarefied = NULL)
sectors = c(4,8,12,16,20,24,28,40,60,80,100)
rotations = c(1,2,3,4,5,10,15,20)

for (sec in sectors){
  for (rot in rotations){
    print(paste0("Sectors: ", sec, ", rotations: ", rot))
    dataset <- optim_list %>% filter(s == sec, r == rot)
    Jump_groups <- group_jumps(dataset, gap_size = 15)
    Jumps_unique <- rarefy_groups(Jump_groups) %>% add_column(Rarefied = TRUE)
    Jumps_unique %<>% select("latitude_rounded", "longitude_rounded", "bio_year", "s", "r", "Rarefied")
    rarefied_list <- rbind(rarefied_list, Jumps_unique)
  }
}


rarefied_list %>% group_by(s, r) %>% summarise(jumps = n())

full_list <- merge(optim_list, rarefied_list, by = c("latitude_rounded", "longitude_rounded", "bio_year", "r", "s"),
                             all = T)
full_list %<>% mutate(Rarefied = ifelse(is.na(Rarefied), "Full", "Rarefied"))

full_list$Rarefied <- factor(full_list$Rarefied, levels = c( "Full", "Rarefied"))

Optim_year <- full_list %>% group_by(sectors = s, rotations = r, bio_year = bio_year, Rarefied = Rarefied) %>% summarise(jumps = n())

head(Optim_year)


#Plot it
optim_plot <- ggplot(data = Optim_year, 
                     aes(x = as.factor(rotations), y = jumps)) +
  geom_bar(lwd = 0.05, stat = "identity", 
           aes(fill = forcats::fct_rev(as.factor(bio_year)),
               alpha = Rarefied)) +
  scale_fill_manual(values = c("#009E73", "#0072B2", "firebrick3", "gold2", "black")) +
  scale_alpha_manual(values = c(0.5, 1)) +
  xlab("Number of rotations") +
  ylab("Number of jumps") +
  facet_wrap(~as.factor(sectors), nrow = 1) +
  guides(fill = guide_legend("Biological year"), alpha = guide_legend("Dataset")) +
  theme_classic() +
  ggtitle("Number of sectors") +
  theme(legend.position = "right", text = element_text(size = 12), axis.text.x = element_text(size = 5, angle = 90), 
        plot.title = NULL)
        #element_text(size = 12, hjust = 0.5))

optim_plot

ggsave(file.path(here(), "figures", "manuscript figures", "S2. optim_plot.jpg"), optim_plot, height = 6, width = 12)
```


Count how long it takes for min number of sectors x rotations
```{r how long does it take?}
system.time({slfdata <- attribute_sectors(grid_data, nb_sectors = 20, centroid = centroid, rotation = 3)
  slfdata_long <- slfdata %>%
  pivot_longer(cols = starts_with("rotation"), names_to = "rotation_nb", values_to = "sectors_nb", names_prefix = "rotation", names_transform = list(rotation_nb = as.integer))
  Results <- find_jumpstest(slfdata_long, gap_size = 15, bio_year = c(2014:2020))})
# 25.79 sec
dim(Results$Jump)[1] #135

# Compare to higher number of sectors but no rotation
system.time({slfdata <- attribute_sectors(grid_data, nb_sectors = 60, centroid = centroid, rotation = 1)
  slfdata_long <- slfdata %>%
  pivot_longer(cols = starts_with("rotation"), names_to = "rotation_nb", values_to = "sectors_nb", names_prefix = "rotation", names_transform = list(rotation_nb = as.integer))
  Results <- find_jumps(slfdata_long, gap_size = 15, bio_year = c(2014:2020))})
# 59.74 sec
dim(Results$Jump)[1] #135
```



# Figure 3: Jump list

Facet 1: map
```{r load dataset, eval = F}

centroid <- c(-75.675340, 40.415240)
Jumps_fig1 <- Jumps %>% mutate(Rarefied = recode(Rarefied,
                                                 "Rarefied" = "Isolated",
                                                 "Full" = "Clustered"))

# Map it
map_rarified <- ggplot(data = grid_data) +
  geom_sf(data = states, fill = "white") +
  geom_point(data = grid_data %>% filter(slf_established == TRUE), aes(x = longitude_rounded, y = latitude_rounded), col = "lightgrey") +
    geom_sf(data = states, alpha = 0) +
    geom_point(data = grid_data, aes(x = centroid[1], y = centroid[2]), col = "black", shape = 4, size = 5) +
  coord_sf(xlim = c(-81, -73), ylim = c(38, 43)) +
  geom_point(data = Jumps_fig1,
             aes(x = longitude_rounded, y = latitude_rounded, 
                 col = as.factor(bio_year), group = Rarefied, shape = Rarefied), stroke = 2, size = 3) +
  scale_discrete_manual(aesthetics = "stroke", values = c('Isolated' = 1, 'Clustered' = 1)) +
  scale_discrete_manual(aesthetics = "shape", values = c('Isolated' = 21, 'Clustered' = 19)) +
  scale_color_manual(values = c("#F0E442", "#D55E00", "#0072B2", "black")) +
  scale_fill_manual(values = c("#F0E442", "#D55E00", "#0072B2", "black")) +
  # scale_stroke_manual(values = c(1, 2)) +
  labs(x = "Longitude", y = "Latitude") +
  theme(legend.position="right", text = element_text(size = 10),
        panel.background = element_rect(fill = "white"),
        legend.key = element_rect(fill = "white")) +
   guides(shape = guide_legend("Jumps"), stroke = guide_legend("Jumps"), colour = guide_legend("Year"), alpha = guide_legend("Jumps"), fill = guide_legend("Year"))

map_rarified

ggsave(file.path(here(), "figures", "manuscript figures", "3. jumps_map_v3.jpg"), map_rarified, height = 8, width = 8)
```

Facet 2: number of jumps per year
```{r number of jumps per year}

Jumps_fig1 <- Jumps %>% mutate(Rarefied = recode(Rarefied,
                                                 "Rarefied" = "Isolated",
                                                 "Full" = "Clustered"))

Jumps_year <- Jumps_fig1 %>% group_by(bio_year, Rarefied) %>% summarise(n = n())
Jumps_fig1$Rarefied <- factor(Jumps_fig1$Rarefied, levels = c( "Clustered", "Isolated"))
 
jumps_plot <- ggplot() +
  geom_bar(data = Jumps_year, aes(x = bio_year, y = n, fill = as.factor(bio_year), group = Rarefied, col = as.factor(bio_year), alpha = Rarefied), stat = "identity", lwd = .25, show.legend = F) +
  scale_fill_manual(values = c("#F0E442", "#D55E00", "#0072B2", "black")) +
  scale_alpha_manual(values = c(1, 0)) +
  scale_color_manual(values = c("#F0E442", "#D55E00", "#0072B2", "black")) +
  xlab("Year") +
  ylab("Number of jumps") +
  theme_classic() +
  guides(alpha = guide_legend("Jumps"), fill = "none", col = "none") +
  theme(text = element_text(size = 10), legend.position = "right")

jumps_plot

ggsave(file.path(here(), "figures", "manuscript figures", "3. number of jumps_v3.jpg"), jumps_plot, height = 2.5, width = 4)
```


Facet 3: mean dist to the threshold per year
```{r dist to threshold per year}
# run and save Results$Dist
# Thresholds <- read.csv("./exported-data/thresholds.csv")
Thresholds %<>% filter(rotation_nb == 1) %>% 
  dplyr::select(longitude_rounded, latitude_rounded, bio_year, sectors_nb) 
head(Thresholds)

Thresholds %<>% 
  dplyr::rename(sectors = sectors_nb,
         latitude_threshold = latitude_rounded,
         longitude_threshold = longitude_rounded)
head(Thresholds)

# Attribute sectors to each jump: find sector for rotation 0
head(Jumps)
class(Jumps)
Jumps$Rarefied <- factor(Jumps$Rarefied, levels = c( "Rarefied", "Full"))

# Jumps %>% filter(Rarefied == "Rarefied")
Jumps_sectors <- attribute_sectors(Jumps, nb_sectors = 20, centroid = c(-75.675340, 40.415240), rotation = 1) 
Jumps_sectors %<>% dplyr::select(-c(theta.1, sectors)) %>%
  dplyr::rename(sectors = rotation1)
Jumps_sectors %>% filter(Rarefied == "Rarefied")
dim(Jumps_sectors)

# Find corresponding threshold per bio year 
JumpLength <- left_join(Jumps_sectors, Thresholds)
JumpLength %<>% rowwise() %>%  mutate(DistToThreshold = as.vector(geosphere::distm(c(longitude_rounded, latitude_rounded), c(longitude_threshold, latitude_threshold), fun = distGeo))/1000)

names(JumpLength)   
head(JumpLength)
JumpLength_tot <- rbind(JumpLength %>% filter(Rarefied == "Rarefied"),
                        JumpLength %>% mutate(Rarefied = "Full"))

# Stats
table(JumpLength_tot$Rarefied)
max(JumpLength_tot$DistToThreshold) #308.90 km

# Full dataset
mean(JumpLength_tot$DistToThreshold) #92.15
quantile(JumpLength_tot$DistToThreshold, 0.95) #240.44 km

# Rare dataset
mean(JumpLength %>% filter(Rarefied == "Rarefied") %>% pull(DistToThreshold)) #87.56
quantile(JumpLength %>% filter(Rarefied == "Rarefied") %>% pull(DistToThreshold), 0.95) #262.55 km


#Boxplots
JumpLength_tot$Rarefied <- factor(JumpLength_tot$Rarefied, levels = c("Rarefied", "Full"))

MeanDist_jump <- ggplot() +
  geom_boxplot(data = JumpLength,
             aes(x = Rarefied, y = DistToThreshold, 
                 col = as.factor(bio_year),
                 fill = as.factor(bio_year)),
                 alpha = 0.7) +
  # scale_alpha_manual(values = c(0.7,0)) +
  scale_color_manual(values = c("#F0E442", "#D55E00", "#0072B2", "black")) +
  scale_fill_manual(values = c("#F0E442", "#D55E00", "#0072B2", "black")) +
  labs(x = "Dataset", y = "Distance from the invasion front (km)") +
  theme_classic() +
  theme(legend.position="top", text = element_text(size = 10), plot.tag.position = c(0.01, 1)) +
  guides(fill = "none", shape = "none", col = "none", alpha = "none")

MeanDist_jump

ggsave(file.path(here(), "figures", "manuscript figures", "3. jump distances_v2.jpg"), MeanDist_jump, height = 2.5, width = 4)
```

Stats
```{r stats jumps distances}
JumpLength_tot %>% group_by(Rarefied, bio_year) %>% 
  summarise(mean(DistToThreshold))

# Full dataset
summary(aov(DistToThreshold ~ bio_year, data = JumpLength)) #p = 0.37

# Rarefied dataset
summary(aov(DistToThreshold ~ bio_year, data = JumpLength %>% filter(Rarefied == "Rarefied"))) #p = 0.34

```


Assemble facets
```{r figure 3 complete}

fig2 <- ggdraw() +
  draw_plot(map_rarified, x = 0, y = .33, width = 1, height = .66) +
  draw_plot(jumps_plot, 0, 0, .5, .33) +
  draw_plot(MeanDist_jump, .5, 0, .5, .33) +
  draw_plot_label(c("(a)", "(b)", "(c)"), c(0, 0, 0.5), c(1, 0.35, 0.35), size = 15) +
  theme(plot.background = element_rect(fill="#FFFFFF", color = NA))

fig2

ggsave(file.path(here(), "figures", "manuscript figures", "3. jump description_v3.jpg"), fig2, height = 10, width = 10)

```




# Other figures (not run)


## Thresholds
Maps of thresholds (all years)
```{r map of threshold points per year per rotation, fig.cap = "Map of SLF jumps", fig.height=6, fig.width=10, eval = F}

centroid <- c(-75.675340, 40.415240)

map_thresholds <- ggplot(data = states) +
  geom_point(data = grid_data %>% filter(slf_established == TRUE),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
   geom_point(data = Results$Dist,
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  facet_wrap(~bio_year, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) +
coord_sf(xlim = c(-82, -72), ylim = c(38, 43), expand = FALSE) +
  theme(legend.position="top")

# ggsave("./figures/vignette_quadrants/map_thresholds.jpg", map_thresholds, width = 15, height = 10)

map_thresholds
```

Map of thresholds, per year
```{r threhsolds per year}
map_thresholds2016 <- ggplot(data = states) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2016)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
   geom_point(data = Results$Dist %>% filter(bio_year == 2016), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  # facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-77, -75), ylim = c(40, 41), expand = FALSE) + 
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_thresholds_2016.jpg", map_thresholds2016 , width = 15, height = 10)


map_thresholds2017 <- ggplot(data = states) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2017)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
   geom_point(data = Results$Dist %>%  filter(bio_year == 2017), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  # facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-77, -74), ylim = c(39.5, 41), expand = FALSE) +  
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_thresholds_2017.jpg", map_thresholds2017 , width = 15, height = 10)


map_thresholds2018 <- ggplot(data = states) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2018)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
   geom_point(data = Results$Dist %>%  filter(bio_year == 2018), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-79, -74), ylim = c(39, 42), expand = FALSE) +
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_thresholds_2018.jpg", map_thresholds2018 , width = 15, height = 10)



map_thresholds2019 <- ggplot(data = states) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2019)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
   geom_point(data = Results$Dist %>%  filter(bio_year == 2019), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-81, -73), ylim = c(39, 42), expand = FALSE) +
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_thresholds_2019.jpg", map_thresholds2019 , width = 15, height = 10)



map_thresholds2020 <- ggplot(data = states) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2020)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
   geom_point(data = Results$Dist %>%  filter(bio_year == 2020), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-82, -72), ylim = c(38, 43), expand = FALSE) + 
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_thresholds_2020.jpg", map_thresholds2020 , width = 15, height = 10)

```


## Jumps
Maps of jumps for each parameter sets.

```{r map of jumps per params per year, fig.cap = "Map of SLF jumps", fig.height=6, fig.width=10, eval = F}

map_jumps2016 <- ggplot(data = states) +
  geom_point(data = Jumps %>%  filter(bio_year %in% c(2014:2015)),
             aes(x = longitude_rounded, y = latitude_rounded), col = "yellow", shape = 19, size = 2) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2016)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
   geom_point(data = Thresholds %>%  filter(bio_year == 2016), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  geom_point(data = Jumps %>%  filter(bio_year == 2016), 
             aes(x = longitude_rounded, y = latitude_rounded), fill = "blue", shape = 21, size = 2) +
  facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-77, -75), ylim = c(40, 41), expand = FALSE) + 
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_jumps_2016.jpg", map_jumps2016 , width = 15, height = 10)


map_jumps2017 <- ggplot(data = states) +
  geom_point(data = Jumps %>%  filter(bio_year %in% c(2014:2016)),
             aes(x = longitude_rounded, y = latitude_rounded), col = "yellow", shape = 19, size = 2) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2017)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data %>%  filter(bio_year == 2017),
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
   geom_point(data = Thresholds %>%  filter(bio_year == 2017), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  geom_point(data = Jumps %>% filter(bio_year == 2017), 
             aes(x = longitude_rounded, y = latitude_rounded), fill = "blue", shape = 21, size = 2) +
  facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-77, -74), ylim = c(39.5, 41), expand = FALSE) + 
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_jumps_2017.jpg", map_jumps2017 , width = 15, height = 10)


map_jumps2018 <- ggplot(data = states) +
  geom_point(data = Jumps %>%  filter(bio_year %in% c(2014:2017)),
             aes(x = longitude_rounded, y = latitude_rounded), col = "yellow", shape = 19, size = 2) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2018)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
   geom_point(data = Thresholds %>%  filter(bio_year == 2018), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  geom_point(data = Jumps %>%  filter(bio_year == 2018), 
             aes(x = longitude_rounded, y = latitude_rounded), fill = "blue", shape = 21, size = 2) +
  facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-79, -74), ylim = c(39, 42), expand = FALSE) + 
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_jumps_2018.jpg", map_jumps2018 , width = 15, height = 10)



map_jumps2019 <- ggplot(data = states) +
  geom_point(data = Jumps %>%  filter(bio_year %in% c(2014:2018)),
             aes(x = longitude_rounded, y = latitude_rounded), col = "yellow", shape = 19, size = 2) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2019)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
  geom_point(data = Thresholds %>%  filter(bio_year == 2019), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  geom_point(data = Jumps %>%  filter(bio_year == 2019), 
             aes(x = longitude_rounded, y = latitude_rounded), fill = "blue", shape = 21, size = 2) +
  facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-81, -73), ylim = c(39, 42), expand = FALSE) + 
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_jumps_2019.jpg", map_jumps2019 , width = 15, height = 10)


map_jumps2020 <- ggplot(data = states) +
  geom_point(data = Jumps %>%  filter(bio_year %in% c(2014:2019)),
             aes(x = longitude_rounded, y = latitude_rounded), col = "yellow", shape = 19, size = 2) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2020)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
  geom_point(data = Thresholds %>%  filter(bio_year == 2020), 
             aes(x = longitude_rounded, y = latitude_rounded), col = "black", shape = 21, size = 2) +
  geom_point(data = Jumps %>%  filter(bio_year == 2020), 
             aes(x = longitude_rounded, y = latitude_rounded), fill = "blue", shape = 21, size = 2) +
  facet_wrap(~params, ncol = 4) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-82, -72), ylim = c(38, 43), expand = FALSE) + 
  theme(legend.position="top")

ggsave("../figures/vignette_quadrants/map_jumps_2020.jpg", map_jumps2020 , width = 15, height = 10)

```

All jumps
```{r map all jumps per params set, eval = F}

# All jumps one set
map_jumps <- ggplot(data = states) +
  geom_point(data = grid_data %>% filter(Status == "Established", bio_year %in% c(2014:2020)),
             aes(x = longitude_rounded, y = latitude_rounded), size = 1, shape = 19, col = "grey") +
  geom_point(data = grid_data,
             aes(x = centroid[1], y = centroid[2]), col = "blue", shape = 4, size = 5) +
  geom_point(data = jumps, 
             aes(x = longitude_rounded, y = latitude_rounded, fill = as.factor(bio_year)), shape = 21, size = 2) +
  geom_text(data = states,
            aes(X, Y, label = code), size = 4) +
  labs(x = "Longitude", y = "Latitude")+
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-82, -72), ylim = c(38, 43), expand = FALSE) + 
  theme(legend.position="top")

ggsave("./figures/vignette_quadrants/map_jumps.jpg", map_jumps , width = 8, height = 8)


```


## Spread distances

```{r figure total spread/diffusive spread for slf established, fig.cap = "Evolution of the radius of the invasion over time, when diffusive spread and jump dispersal are separated", echo=FALSE, message = FALSE, eval = F}

#Data on total spread

spread_distances_tab <- grid_data_rotate0 %>% filter(Status == "Established" & bio_year %in% c(2014:2020)) %>%
  group_by(Status, bio_year, portion) %>% 
  summarise(MaxDist = max(DistToIntro))


spread <- ggplot(data = spread_distances_tab, aes(x = bio_year, y = MaxDist)) +
  facet_wrap(~portion, ncol = 4) +
    geom_bar(stat="identity", fill = "white", col = "black") +
  geom_bar(data = Results_rotate4$Dist, aes(x = bio_year , y = DistToIntro), fill = "grey", col = "black", stat = "identity") +
  ylab("Invasion radius (km)")+
  xlab("Year") +
  theme_classic()

spread
```



# Yearly radius of the invasion


To estimate the spread of the SLF, we extract for each year the radius of the invasion, defined as the maximum distance of a survey with established SLF to the introduction point. Median or mean distances are not informative here because surveys are preferentially conducted towards the invasion front, and thus bias the distribution of distances.

```{r advance of the invasion, fig.cap = "Yearly radius of the invasion", fig.height = 5, fig.width = 9, warning = F, echo = params$display, message = FALSE}

#Computes the maximum distance to the introduction site for each year, status and quadrant
advance_tab <- grid_data_DP12_rotate0 %>% filter(Status == "Established" & bio_year != 2021) %>%
  group_by(Status, bio_year, portion_name) %>% 
  summarise(MaxDist = max(DistToIntro))

# same thing but in a figure
advance_graph <- ggplot(data = advance_tab, aes(x = bio_year , y = MaxDist)) +
  geom_bar(stat="identity", fill = "darkgrey", col = "black") +
  ylab("Invasion radius (km)")+
  # ylim(c(0,1000)) +
  xlab("Year") +
  facet_wrap(~portion_name, ncol = 4) +
  theme_classic()

advance_graph

```

```{r save graph advance radius, eval = params$savefiles, echo = params$display}
ggsave("../figures/vignette_quadrants/advance_graph.jpg", advance_graph, width = 8, height = 6)
```

The invasion radius increased regularly until 2017 (Figure 2). From 2018, there is a steep increase in the invasion radius, especially in the westernmost disk portions, denoting the apparition of dispersal jumps. Note that in 2020, the invasion radius did not increase dramatically, and even decreased in the NNE and ESE portions. This might be due to fewer surveys being conducted that year because of the covid19 pandemic, or be an actual biological pattern. The accuracy of these values can be investigated by checking whether negative surveys are found further than positive surveys, to make sure that SLF were not missed.


## Sectors
```{r test run function to attribute sectors, echo = params$display, eval = params$run}

# Reload the grid_data file directly
grid_data <- read.csv(file.path(here(), "exported-data", "grid_data.csv"), h=T)
centroid <- c(-75.675340, 40.415240)

# Examples
# 8 sectors
grid_data_8S <- attribute_sectors(grid_data, nb_sectors = 8, centroid = centroid, rotation = 10)
unique(grid_data_8S$bio_year)

# 20 sectors
grid_data_20S <- attribute_sectors(grid_data, nb_sectors = 20, centroid = centroid, rotation = 10)

# Bind datasets together
grid_data_allrotations_S8_20 <- bind_rows(grid_data_8S, grid_data_20S)

centroid <- c(long = -75.675340, lat = 40.415240)

#pivot_longer for figures
grid_data_long <- grid_data_allrotations_S8_20 %>% 
  pivot_longer(cols = starts_with("rotation"), names_to = "rotation_nb", values_to = "sectors_nb", names_prefix = "rotation", names_transform = list(rotation_nb = as.integer)) 
```

```{r map of 8-20 sectors together, fig.cap = "Maps of the different number of sectors used, with established SLF colored by sector (2014-2020)", fig.height = 6, fig.width = 8, warning = F, echo = params$display}


# map_8_20S <-  ggplot(data = states) +
#   # geom_point(data = grid_data_long %>% filter(Status == "Established" & bio_year %in% c(2014:2020)),
#              # aes(x = longitude_rounded, y = latitude_rounded, col = as.factor(sectors_nb)),
#              # shape = 19, size = 2, show.legend=F) +
#   # scale_color_manual(values=c(rep(c("lightblue", "yellow", "green", "pink"), 5))) +
#   # geom_point(data = grid_data_long,
#              # aes(x = centroid[1], 
#                  # y = centroid[2]),
#              # col = "blue", shape = 4, size = 5) +
#   # geom_text(data = states, aes(X, Y, label = code), size = 4.5) +
#   geom_sf(data = states, alpha = 0) + 
#   # facet_wrap(~sectors, ncol = 4) +
#   coord_sf(xlim = c(-81, -73), ylim = c(38, 42.5), expand = FALSE) +
#   labs(x = "Longitude", y = "Latitude", col = "Sectors")
# 
# 
# map_8_20S_all <-  ggplot(data = states) +
#   geom_point(data = grid_data_long %>% filter(sectors == 8),
#              aes(x = longitude_rounded, y = latitude_rounded, col = as.factor(sectors_nb)), 
#              shape = 19, size = 2, show.legend=F) +
#   scale_color_manual(values=c(rep(c("lightblue", "yellow", "green", "pink"), 5))) +
#   geom_point(data = grid_data_long,
#              aes(x = centroid[1], 
#                  y = centroid[2]),
#              col = "blue", shape = 4, size = 5) +
#   # geom_text(data = states, aes(X, Y, label = code), size = 4.5) +
#   geom_sf(data = states, alpha = 0) + 
#   facet_wrap(~sectors, ncol = 4) +
#   coord_sf(xlim = c(-81, -73), ylim = c(38, 42.5)) +
#   labs(x = "Longitude", y = "Latitude", col = "Sectors")

```


```{r map of 8-12-16 sectors separated with 0 rotation, fig.cap = "Map of surveys with established SLF colored by sector (2014-2020)", fig.height = 6, fig.width = 8, warning = F, echo = params$display}

map_8sectors <-  ggplot(data = states) +
  geom_point(data = grid_data_long %>% filter(sectors == 8 & Status == "Established"),
             aes(x = longitude_rounded, y = latitude_rounded, col = as.factor(sectors_nb)), 
             shape = 19, size = 2, show.legend=F) +
  scale_color_manual(values=c(rep(c("lightblue", "yellow", "green", "pink"), 4))) +
  geom_point(data = grid_data_long %>% filter(sectors == 8 & Status == "Established"),
             aes(x = centroid[1], 
                 y = centroid[2]),
             col = "blue", shape = 4, size = 5) +
  geom_text(data = states, aes(X, Y, label = code), size = 2) +
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-90, -70), ylim = c(35, 45)) +
  labs(x = "Longitude", y = "Latitude", col = "Sectors")

ggsave("../figures/vignette_quadrants/sectors_map/map_8sectors_rotate0.jpg", map_8sectors, width = 6, height = 6)


map_20sectors <-  ggplot(data = states) +
  geom_point(data = grid_data_long %>% filter(sectors == 20 & Status == "Established"),
             aes(x = longitude_rounded, y = latitude_rounded, col = as.factor(sectors_nb)), 
             shape = 19, size = 2, show.legend=F) +
  scale_color_manual(values=c(rep(c("lightblue", "yellow", "green", "pink"), 5))) +
  geom_point(data = grid_data_long %>% filter(sectors == 20 & Status == "Established"),
             aes(x = centroid[1], 
                 y = centroid[2]),
             col = "blue", shape = 4, size = 5) +
  geom_text(data = states, aes(X, Y, label = code), size = 2) +
  geom_sf(data = states, alpha = 0) + 
  coord_sf(xlim = c(-90, -70), ylim = c(35, 45)) +
  labs(x = "Longitude", y = "Latitude", col = "Sectors")

ggsave("../figures/vignette_quadrants/sectors_map/map_20sectors_rotate0.jpg", map_20sectors, width = 6, height = 6)
```


## Rotations
```{r map of each rotation of sectors for each number of sectors, fig.cap = "Map of surveys with established SLF colored by sector (2014-2020)", fig.height = 6, fig.width = 8, warning = F, echo = params$display}

map_8sectors_wrap <-  ggplot(data = states) +
  geom_point(data = grid_data_long %>% filter(sectors == 8 & Status == "Established"),
             aes(x = longitude_rounded, y = latitude_rounded, col = as.factor(sectors_nb)), 
             shape = 19, size = 2, show.legend=F) +
  scale_color_manual(values=c(rep(c("lightblue", "yellow", "green", "pink"), 2))) +
  geom_point(data = grid_data_long %>% filter(sectors == 8 & Status == "Established"),
             aes(x = centroid[1], 
                 y = centroid[2]),
             col = "blue", shape = 4, size = 5) +
  geom_text(data = states, aes(X, Y, label = code), size = 2) +
  geom_sf(data = states, alpha = 0) + 
  facet_wrap(~rotation_nb, ncol = 5) +
  coord_sf(xlim = c(-81, -73), ylim = c(38, 42.5), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude", col = "Sectors")

ggsave("../figures/vignette_quadrants/sectors_map/map_8sectors.jpg", map_8sectors_wrap, width = 15, height = 6)

map_20sectors_wrap <-  ggplot(data = states) +
  geom_point(data = grid_data_long  %>% filter(sectors == 20 & Status == "Established"),
             aes(x = longitude_rounded, y = latitude_rounded, col = as.factor(sectors_nb)), 
             shape = 19, size = 2, show.legend=F) +
  scale_color_manual(values=c(rep(c("lightblue", "yellow", "green", "pink"), 5))) +
  geom_point(data = grid_data_long %>% filter(sectors == 20 & Status == "Established"),
             aes(x = centroid[1], 
                 y = centroid[2]),
             col = "blue", shape = 4, size = 5) +
  geom_text(data = states, aes(X, Y, label = code), size = 2) +
  geom_sf(data = states, alpha = 0) + 
  facet_wrap(~rotation_nb, ncol = 5) +
  coord_sf(xlim = c(-81, -73), ylim = c(38, 42.5), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude", col = "Sectors")

ggsave("../figures/vignette_quadrants/sectors_map/map_20sectors.jpg", map_20sectors_wrap, width = 15, height = 6)


```
