---
title: "Untitled"
author: "Nadege Belouard"
date: "1/18/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(spData)
library(RVAideMemoire)
library(dplyr)
library(gridExtra)
library(magrittr)
library(purrr)
library(here)
library(FSA)

```

# 7. Generate random dispersal distribution


Generate all points
```{r prepare set of possible points, eval = T}

# Map slfEStab
grid <- read.csv(file.path(here(), "exported-data", "grid_data.csv"))
head(grid_data)
grid <- grid_data
dim(grid_data %>% filter(bio_year != 2021))
dim(grid_data)
# slfEstab <- grid %>% filter(Status == "Established", bio_year != 2021)
slfEstab <- grid_data %>% filter(slf_established == TRUE)
slfEstab_layer <- st_as_sf(x = slfEstab, coords = c("longitude_rounded", "latitude_rounded"), crs = "EPSG:4269", remove = F)
slfEstab_layer <- st_transform(slfEstab_layer, crs = "ESRI:102010")

US <- st_read("C:/Users/labuser/Documents/Postdoc_SLF/SLF_Dispersal/data/raw_data/states/gadm36_Cont_USA_county/gadm36_Cont_USA_county.shp", crs = "EPSG:4326", quiet = T)
st_crs(US)
US <- st_transform(US, crs = "ESRI:102010")
Pennsylvania <- US %>% filter(NAME_1 == "Pennsylvania")

# Visualize the SLF data
ggplot(data = US, fill = "white") +
  geom_sf() +
  geom_sf(data = slfEstab_layer, col = "blue") +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)



## Generate all possible coordinates of the random distribution
# Calculate all potential coordinates
maxlat <- max(slfEstab$latitude_rounded)
minlat <- min(slfEstab$latitude_rounded)
maxlong <- max(slfEstab$longitude_rounded)  
minlong <- min(slfEstab$longitude_rounded)

Seqlat <- seq(from = minlat, to = maxlat, by = 1/111)
Seqlong <- seq(from = minlong, to = maxlong, by = 1/85)
Coordinates <- expand.grid(latitude = Seqlat, longitude = Seqlong)
Coordinates <- st_as_sf(x = Coordinates, coords = c("longitude", "latitude"), crs = "EPSG:4269", remove = F)
Coordinates <- st_transform(Coordinates, crs = "ESRI:102010")

ggplot(data = US, fill = "white") +
  geom_sf() +
  geom_sf(data = Coordinates, alpha = 0.5) +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)


# Create the minimum convex polygon of the SLF data
hull <- st_convex_hull(st_union(slfEstab_layer))

chull <- ggplot(data = states, fill = "white") +
  geom_sf() +
  # geom_sf(data = Coordinates, alpha = 0.5) +
  geom_sf(data = hull, alpha = 0.5, fill = "yellow") + 
  # geom_sf(data = buffer, alpha = 0.5, fill = "red") +
  # geom_sf(data = slfEstab_layer, col = "blue") +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)


# Keep only the coordinates that are in the minimum convex polygon
## The alternative is to keep all coordinates that are in a buffer around the SLF detections
Coordinates_chull <- st_intersection(Coordinates, hull)

ggplot(data = US, fill = "white") +
  geom_sf() +
  geom_sf(data = Coordinates_chull, alpha = 0.5) +
  geom_sf(data = slfEstab_layer, col = "blue") +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)


#Keep only those that are on land!
Coordinates_land <- st_intersection(Coordinates_chull, US)

random_coord <- ggplot(data = states, fill = "white") +
  geom_sf() +
  # geom_sf(data = Coordinates_land, alpha = 0.5) +
  geom_sf(data = hull, alpha = 0.5, fill = "white") + 
  geom_point(data = slfEstab_layer,
             aes(x = longitude_rounded, y = latitude_rounded), col = "black") +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(-82, -72), ylim = c(37, 43)) +
  theme_classic()

Coordinates_land %<>% select(latitude, longitude, geometry)
st_write(Coordinates_land, file.path(here(), "exported-data", "Distances_random distrib", "coordinates_set.shp"), driver = "ESRI Shapefile")

Coordinates_land <- st_read(file.path(here(), "exported-data", "Distances_random distrib", "coordinates_set.shp"), quiet = T)





# Keep only surveys that are within the chull
slf_layer <- st_as_sf(x = slf_uptodate, coords = c("longitude_rounded", "latitude_rounded"), crs = "EPSG:4269", remove = F)
slf_layer <- st_transform(slf_layer, crs = "ESRI:102010")
slf_chull <- st_intersection(slf_layer, hull)

ggplot(data = states, fill = "white") +
  geom_sf() +
  geom_sf(data = slf_layer) +
  geom_sf(data = hull, alpha = 0.5, fill = "yellow") + 
  geom_sf(data = slf_chull, col = "blue") +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)

st_geometry(slf_chull) <- NULL
write.csv(slf_chull, file.path(here(), "exported-data", "points_chull.csv"), row.names = F)

slf_chull <- read.csv(file.path(here(), "exported-data", "points_chull.csv"))
dim(slf_chull %>% filter(Status == "Positive surveys"))[1]/dim(slf_chull)[1] #32,743 points
table(slf_chull$Status) #18% positive points

```


Calculate distances for the grid of random points (step done on the HPC - computationally intensive)
```{r sample points and calculate distances}
dim(jumpers)
Coordinates_land <- st_read(file.path(here(), "exported-data", "Distances_random distrib", "coordinates_set.shp"), quiet = T)


# Create rows for distances
Coordinates_land %<>% add_column(DistToRail = NA,
                          DistToRoad = NA,
                          DistToAirport = NA,
                          DistToPort = NA)

#Calculate their distance to transport infrastructures
for (j in 1:length(Coordinates_land$DistToRail)){ 
  # Print the row being considered
  print(j)
  
  point_clip <- st_buffer(Coordinates_land[j,], dist = 50000)
  
  # Calculate distance to the closest railroad
  rail_clip <- st_intersection(rail, point_clip)
  dist_rail <- st_distance(x = Coordinates_land[j,], y = rail_clip)
  Coordinates_land$DistToRail[j] <- min(dist_rail)
  
  # Calculate distance to the closest major road
  road_clip <- st_intersection(road, point_clip)
  dist_road <- st_distance(x = Coordinates_land[j,], y = road_clip)
  Coordinates_land$DistToRoad[j] <- min(dist_road)
  
  # Calculate distance to the closest international/millitary airport
  dist_airport <- st_distance(x = Coordinates_land[j,], y = airports)
  Coordinates_land$DistToAirport[j] <- min(dist_airport)
  
  # Calculate distance to the closest port
  dist_port <- st_distance(x = Coordinates_land[j,], y = ports)
  Coordinates_land$DistToPort[j] <- min(dist_port)
  
  rm(point_clip, rail_clip, road_clip)
  
  # Stores the result every 1,000 iteration to avoid loosing all the data in case of crash
  if (i %% 1000 == 0){
    st_write(Coordinates_land, "./exported-data/Distances_random distrib/random_distances_transport.shp", row.names = F)
  }
}


# Save file
st_geometry(Coordinates_land) <- NULL
write.csv(Coordinates_land, "./exported-data/Distances_observed/random_distances_transport.csv", row.names = F)
```


The dataset was computed in 4 files on the HPC, we combine them in one file.

```{r assemble the dataset}
random_dist1 <- read.csv(file.path(here(), "exported-data", "Distances_random distrib", "random_distances_transport0-66000.csv"))
random_dist2 <- read.csv(file.path(here(), "exported-data", "Distances_random distrib", "random_distances_transport66-100000.csv"))
random_dist3 <- read.csv(file.path(here(), "exported-data", "Distances_random distrib", "random_distances_transport100-140000.csv"))
random_dist4 <- read.csv(file.path(here(), "exported-data", "Distances_random distrib", "random_distances_transport140-end.csv"))

Coordinates_land <- rbind(random_dist1[0:66000,], random_dist2[66001:100000,],
      random_dist3[100001:140000,], random_dist4[140001:length(random_dist4$latitude),])
dim(Coordinates_land) == dim(random_dist1)

# Verify that there is no 0
Coordinates_land %>% filter(DistToRail == Inf | is.na(DistToRail))
Coordinates_land %>% filter(DistToRoad == Inf | is.na(DistToRoad))

write.csv(Coordinates_land, "./exported-data/Distances_random distrib/Random_transport.csv", row.names = F)

```


# Select random datasets

```{r select random datasets}

Coordinates_land <- read.csv(file.path(here(), "exported-data", "Distances_random distrib", "Random_transport.csv"))
head(Coordinates_land)

slf_uptodate <- read.csv(file.path(here(), "exported-data", "slf_uptodate.csv"), h=T)

slf_uptodate <- read.csv(file.path(here(), "exported-data", "points_chull.csv"), h=T)

# Size of the dataset to be sampled
jumpers_full = dim(slf_uptodate %>% filter(Category_full == "Jumpers"))[1]
jumpers_rarefied = dim(slf_uptodate %>% filter(Category_rare == "Jumpers"))[1]
diffusers_full = dim(slf_uptodate %>% filter(Category_full == "Diffusers"))[1]
diffusers_rarefied = dim(slf_uptodate %>% filter(Category_rare == "Diffusers"))[1]
negatives = dim(slf_uptodate %>% filter(Category_full == "Negatives"))[1]
#There is no difference in the negatives between the full and rarefied datasets!


for (i in 1:1){
  #Generate a set of coordinates
  Random_jumpers_full <- Coordinates_land[sample(nrow(Coordinates_land), size = jumpers_full, replace = F),] %>%
    add_column(Category_full = "Jumpers")
  Random_diffusers_full <- Coordinates_land[sample(nrow(Coordinates_land), size = diffusers_full, replace = F),] %>%
    add_column(Category_full = "Diffusers")
  Random_negatives <- Coordinates_land[sample(nrow(Coordinates_land), size = negatives, replace = F),] %>%
    add_column(Category_full = "Negatives")
  
  Random_coordinates <- rbind(Random_jumpers_full, Random_diffusers_full,
                              Random_negatives)
  
  #Calculate the mean and median distance per simulation
  Random_means <- Random_coordinates %>% group_by(Category_full) %>% 
    summarise_at(vars(starts_with("DistTo")), list(mean = mean, median = median)) %>% 
    add_column(Simulation = i)
  
  #Save the table with the simulation number
  if (i == 1){
    Simulations <- Random_means
  } else {
    Simulations <- bind_rows(Simulations, Random_means)
  }
  
  if (i %% 100 == 0){ print(i)}
} 

# write.csv(Simulations, "./exported-data/Distances_random distrib/SimulatedMeans_fulluptodate.csv", row.names = F)
write.csv(Simulations, "./exported-data/Distances_random distrib/SimulatedMeans_fullchull.csv", row.names = F)


# Same for rarefied dataset

for (i in 1:9999){
  #Generate a set of coordinates
  Random_jumpers_rarefied <- Coordinates_land[sample(nrow(Coordinates_land), size = jumpers_rarefied, replace = F),] %>%
    add_column(Category_rarefied = "Jumpers")
  Random_diffusers_rarefied <- Coordinates_land[sample(nrow(Coordinates_land), size = diffusers_rarefied, replace = F),] %>%
    add_column(Category_rarefied = "Diffusers")
  Random_negatives <- Coordinates_land[sample(nrow(Coordinates_land), size = negatives, replace = F),] %>%
    add_column(Category_rarefied = "Negatives")
  
  Random_coordinates <- rbind(Random_jumpers_rarefied, Random_diffusers_rarefied,
                              Random_negatives)
  
  #Calculate the mean and median distance per simulation
  Random_means <- Random_coordinates %>% group_by(Category_rarefied) %>% 
    summarise_at(vars(starts_with("DistTo")), list(mean = mean, median = median)) %>% 
    add_column(Simulation = i)
  
  #Save the table with the simulation number
  if (i == 1){
    Simulations <- Random_means
  } else {
    Simulations <- bind_rows(Simulations, Random_means)
  }
  
  if (i %% 100 == 0){ print(i)}
} 

write.csv(Simulations, "./exported-data/Distances_random distrib/SimulatedMeans_rarefiedchull.csv", row.names = F)

```


Map points
```{r map of random points for jumpers, fig.width= 6, fig.height=7, eval = FALSE}
# Check on the first randomly generated dataset that the points location corresponds to the expectation in each disk portion (Figure 2).

# All possible coordinates
ggplot(data = states) +
  geom_sf(fill = "white") + 
  geom_sf(data = Coordinates_land, col = "blue") +
  labs(x = "Longitude", y = "Latitude")+
  theme(legend.position="top")


# Example of random coordinates
ggplot(data = states) +
  geom_sf(fill = "white") + 
  geom_point(data = Random_jumpers_full, aes(x = longitude, y = latitude), col = "blue") +
  labs(x = "Longitude", y = "Latitude")+
  geom_text(data = states, aes(X, Y, label = code), size = 5) +
  theme(legend.position="top") +
  coord_sf(xlim = c(-82, -72), ylim = c(37, 43), expand = FALSE)
  

# Actual jumps
jumpers <- read.csv(file.path(here(), "exported-data", "jumps_full_rarefied.csv"))
jumpers <- st_as_sf(x = jumpers, coords = c("longitude_rounded", "latitude_rounded"), crs = "EPSG:4269", remove = F)
jumpers <- st_transform(jumpers, crs = "ESRI:102010")
st_crs(jumpers)

ggplot(data = states) +
  geom_sf(fill = "white") + 
  geom_sf(data = jumpers, col = "red") +
  labs(x = "Longitude", y = "Latitude")+
  theme(legend.position="top")

```



# Visualize results

slf chull
```{check results with negatives in chull only}

# Load the simulated values
Simulations_rare <- read.csv("./exported-data/Distances_random distrib/SimulatedMeans_rarefiedchull.csv")
Simulations_full <- read.csv("./exported-data/Distances_random distrib/SimulatedMeans_fullchull.csv")

# Convert simulations to long format
Simulations_full_long <- Simulations_full %>% pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")
Simulations_full_long$Category_full <- factor(Simulations_full_long$Category_full, levels = c("Jumpers", "Diffusers", "Negatives"))

Simulations_full_long_mean <- Simulations_full_long %>% filter(DistanceType %in% c("DistToRail_mean", "DistToRoad_mean", "DistToAirport_mean", "DistToPort_mean"))
Simulations_full_long_mean$DistanceType <- as.factor(Simulations_full_long_mean$DistanceType)
levels(Simulations_full_long_mean$DistanceType)
levels(Simulations_full_long_mean$DistanceType) <- c("DistToAirport", "DistToPort", "DistToRail", "DistToRoad")


Simulations_full_long_median <- Simulations_full_long %>% filter(DistanceType %in% c("DistToRail_median", "DistToRoad_median", "DistToAirport_median", "DistToPort_median"))
Simulations_full_long_median$DistanceType <- as.factor(Simulations_full_long_median$DistanceType)
levels(Simulations_full_long_median$DistanceType)
levels(Simulations_full_long_median$DistanceType) <- c("DistToAirport", "DistToPort", "DistToRail", "DistToRoad")


# Convert the observed data to long format, get the average values
slf_uptodate <- read.csv(file.path(here(), "exported-data", "points_chull.csv"))
dim(slf_uptodate) #32,743 rows


# test with only points from the convex hull
# slf_uptodate <- slf_chull
slf_uptodate_long <- slf_uptodate %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")
slf_obsmeans <- slf_uptodate_long %>% 
  group_by(Category_full, DistanceType) %>% 
  summarise(#MeanDistance = mean(DistanceValue),
            MedianDistance = median(DistanceValue))
slf_obsmeans$Category_full <- factor(slf_obsmeans$Category_full, levels = c("Jumpers", "Diffusers", "Negatives"))


# Full dataset
# Plot observed vs simulated means
random_transport <- ggplot() +
  geom_histogram(data = Simulations_full_long_mean, 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_full), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans,
             mapping = aes(xintercept = MeanDistance/1000, col = Category_full), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~Category_full + DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_fullmeanschull.jpg"), random_transport, width = 10, height = 9)




## Jumpers only
random_transport <- ggplot() +
  geom_histogram(data = Simulations_full_long_mean %>% filter(Category_full == "Jumpers"), 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_full), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans %>% filter(Category_full == "Jumpers"),
             mapping = aes(xintercept = MeanDistance/1000, col = Category_full), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_fullmeanschull_jumpers.jpg"), random_transport, width = 10, height = 3)



# Plot observed vs simulated medians
random_transport <- ggplot() +
  geom_histogram(data = Simulations_full_long_median, 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_full), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans,
             mapping = aes(xintercept = MedianDistance/1000, col = Category_full), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~Category_full + DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_fullmedianschull.jpg"), random_transport, width = 10, height = 9)


# Jumpers only
random_transport <- ggplot() +
  geom_histogram(data = Simulations_full_long_median %>% filter(Category_full == "Jumpers"), 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_full), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans %>% filter(Category_full == "Jumpers"),
             mapping = aes(xintercept = MedianDistance/1000, col = Category_full), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

random_transport

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_fullmedianschull_jumpers.jpg"), random_transport, width = 10, height = 3)


# COUNT HOW MANY SIMULATIONS ARE SMALLER THAN THE OBS VALUE
obsrail <- slf_obsmeans %>% filter(Category_full == "Jumpers", DistanceType == "DistToRail") %>% pull(MedianDistance) 
dim(Simulations_full_long_median %>% filter(Category_full == "Jumpers", DistanceType == "DistToRail", DistanceValue < obsrail))[1]
#0

obsroad <- slf_obsmeans %>% filter(Category_full == "Jumpers", DistanceType == "DistToRoad") %>% pull(MedianDistance) 
dim(Simulations_full_long_median %>% filter(Category_full == "Jumpers", DistanceType == "DistToRoad", DistanceValue < obsrail))[1]
#0

obsairport <- slf_obsmeans %>% filter(Category_full == "Jumpers", DistanceType == "DistToAirport") %>% pull(MedianDistance) 
dim(Simulations_full_long_median %>% filter(Category_full == "Jumpers" & DistanceType == "DistToAirport" & DistanceValue < obsairport))[1]
#1728

obsport <- slf_obsmeans %>% filter(Category_full == "Jumpers", DistanceType == "DistToPort") %>% pull(MedianDistance) 
dim(Simulations_full_long_median %>% filter(Category_full == "Jumpers" & DistanceType == "DistToPort" & DistanceValue < obsport))[1]
#9997


# Same with rarefied
# Convert simulations to long format
Simulations_rare_long <- Simulations_rare %>% pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")
Simulations_rare_long %<>% rename(Category_rare = Category_rarefied) 
Simulations_rare_long$Category_rare <- factor(Simulations_rare_long$Category_rare, levels = c("Jumpers", "Diffusers", "Negatives"))

Simulations_rare_long_mean <- Simulations_rare_long %>% filter(DistanceType %in% c("DistToRail_mean", "DistToRoad_mean", "DistToAirport_mean", "DistToPort_mean"))
Simulations_rare_long_mean$DistanceType <- as.factor(Simulations_rare_long_mean$DistanceType)
levels(Simulations_rare_long_mean$DistanceType)
levels(Simulations_rare_long_mean$DistanceType) <- c("DistToAirport", "DistToPort", "DistToRail", "DistToRoad")


Simulations_rare_long_median <- Simulations_rare_long %>% filter(DistanceType %in% c("DistToRail_median", "DistToRoad_median", "DistToAirport_median", "DistToPort_median"))
Simulations_rare_long_median$DistanceType <- as.factor(Simulations_rare_long_median$DistanceType)
levels(Simulations_rare_long_median$DistanceType)
levels(Simulations_rare_long_median$DistanceType) <- c("DistToAirport", "DistToPort", "DistToRail", "DistToRoad")


# Convert the observed data to long format, get the average values
slf_uptodate <- read.csv(file.path(here(), "exported-data", "points_chull.csv"))
dim(slf_uptodate) #32,743 rows

# test with only points from the convex hull
# slf_uptodate <- slf_chull
slf_uptodate_long <- slf_uptodate %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")
slf_obsmeans <- slf_uptodate_long %>% 
  group_by(Category_rare, DistanceType) %>% 
  summarise(MeanDistance = mean(DistanceValue),
            MedianDistance = median(DistanceValue))
slf_obsmeans$Category_rare <- factor(slf_obsmeans$Category_rare, levels = c("Jumpers", "Diffusers", "Negatives"))


# Plot observed vs simulated means
random_transport <- ggplot() +
  geom_histogram(data = Simulations_rare_long_mean, 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_rare), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans,
             mapping = aes(xintercept = MeanDistance/1000, col = Category_rare), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~Category_rare + DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_raremeanschull.jpg"), random_transport, width = 10, height = 9)



# Plot observed vs simulated medians
random_transport <- ggplot() +
  geom_histogram(data = Simulations_rare_long_median, 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_rare), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans,
             mapping = aes(xintercept = MedianDistance/1000, col = Category_rare), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~Category_rare + DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_raremedianschull.jpg"), random_transport, width = 10, height = 9)


## Jumpers
random_transport <- ggplot() +
  geom_histogram(data = Simulations_rare_long_median %>% filter(Category_rare == "Jumpers"), 
                 aes(x =  DistanceValue/1000, y = ..density.., fill = Category_rare), binwidth = 0.1) +
  geom_vline(data = slf_obsmeans %>% filter(Category_rare == "Jumpers"),
             mapping = aes(xintercept = MedianDistance/1000, col = Category_rare), size = 1) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Distance to the nearest... (km)") +
  ylab("Count (simulations)") +
  facet_wrap(~DistanceType, ncol = 4, scale = "free") +
  theme_classic() +
  guides(col = guide_legend("SLF category"), fill = guide_legend("SLF category"))

ggsave(file.path(here(), "figures", "vignette_transports", "bootstrap_transports_raremedianschull_jumpers.jpg"), random_transport, width = 10, height = 3)



# COUNT HOW MANY SIMULATIONS ARE SMALLER THAN THE OBS VALUE
obsrail <- slf_obsmeans %>% filter(Category_rare == "Jumpers", DistanceType == "DistToRail") %>% pull(MedianDistance) 
dim(Simulations_rare_long_median %>% filter(Category_rare == "Jumpers", DistanceType == "DistToRail", DistanceValue < obsrail))[1]
#0

obsroad <- slf_obsmeans %>% filter(Category_rare == "Jumpers", DistanceType == "DistToRoad") %>% pull(MedianDistance) 
dim(Simulations_rare_long_median %>% filter(Category_rare == "Jumpers", DistanceType == "DistToRoad", DistanceValue < obsrail))[1]
#148

obsairport <- slf_obsmeans %>% filter(Category_rare == "Jumpers", DistanceType == "DistToAirport") %>% pull(MedianDistance) 
dim(Simulations_rare_long_median %>% filter(Category_rare == "Jumpers" & DistanceType == "DistToAirport" & DistanceValue < obsairport))[1]
#620

obsport <- slf_obsmeans %>% filter(Category_rare == "Jumpers", DistanceType == "DistToPort") %>% pull(MedianDistance) 
dim(Simulations_rare_long_median %>% filter(Category_rare == "Jumpers" & DistanceType == "DistToPort" & DistanceValue < obsport))[1]
#6647

slf_obsmeans

``` 





```{r check distribution of observed distances}
slf_uptodate <- read.csv(file.path(here(), "exported-data", "points_chull.csv"))
dim(slf_uptodate) #43,801 rows

# slf_uptodate <- slf_chull
slf_uptodate_long <- slf_uptodate %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")

slf_uptodate_long$Category_full <- factor(slf_uptodate_long$Category_full, levels = c("Jumpers", "Diffusers", "Negatives"))

distrib <- ggplot(slf_uptodate_long, aes(x =  DistanceValue/1000, y = ..density.., fill = Category_full), binwidth = 0.1) +
  geom_histogram() +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~Category_full+DistanceType, scales = "free") +
  theme_classic() 


ggsave(file.path(here(), "figures", "vignette_transports", "distribution_observed_chull.jpg"), distrib, width = 6, height = 6)
```


# SLF 101
```{r slf 101}
# Simulated means
MeanDistances <- read.csv(file.path(here(), "exported-data", "Distances_random distrib", "SimulatedMeans_fullchull.csv"))
names(MeanDistances)
head(MeanDistances)

#Modify dataset to get distribution of distances with a column for the type of distance, and one column for the type of dataset (Full or Rarefied)
random_long <- MeanDistances %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")


## FOR MATT SLF 101 MEAN FULL
random_long_mean <- random_long %>% filter(grepl("_mean", DistanceType))
head(random_long_mean)
random_long_mean$DistanceType <- as.factor(random_long_mean$DistanceType)
levels(random_long_mean$DistanceType)
levels(random_long_mean$DistanceType) <- gsub(pattern = "_mean", replacement = "", x = levels(random_long_mean$DistanceType))


head(random_long_mean)
summary_sim_full <- random_long_mean %>% group_by(Category_full, DistanceType) %>% summarise(mean_sim = mean(DistanceValue), sd_sim = sd(DistanceValue))
head(summary_sim_full)


# Observed data
slf_uptodate <- read.csv(file.path(here(), "exported-data", "points_chull.csv"))
slf_cat_long <- slf_uptodate %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")

slf_obsmeans <- slf_cat_long %>% 
  group_by(Category_full, DistanceType) %>% 
  summarise(MeanDistance = mean(DistanceValue)#,
            # MedianDistance = median(DistanceValue)
            )


summary_full <- merge(summary_sim_full, slf_obsmeans)
summary_full %<>% rename(mean_obs = MeanDistance) %>% 
  mutate(effect_size = (mean_obs - mean_sim)/sd_sim)
write.csv(summary_full, "summary_simulations_full_transports.csv")


Category = as.vector(unique(summary_full$DistanceType))
str(Category)
Properties = c("Primary airports (more than 10,000 passenger boardings each year, FAA, USGS National Transportation Dataset)",
               "Ports (USGS Science-Base Catalog",
               "Railroads (USGS National Transportation Dataset)",
               "Major roads (primary and secondary roads as defined by the MAF/TIGER Feature Classification, US Census Bureau)")
Names <- data.frame(DistanceType = Category, Properties = Properties)


summary_full_cat <- merge(summary_full, Names)
write.csv(summary_full_cat, "summary_simulations_full_transports.csv")


# GRAPH
ggplot(summary_full_cat, aes(y = effect_size, x = DistanceType)) + 
  geom_point(aes(col = Category_full)) +
  # scale_fill_brewer(palette = "Dark2") +
  # facet_wrap(~Category_full, scales = "free_y", ncol = 1) +
             # labeller = labeller(DistanceType = 
  #   c("DistToAirport" = "Airport",
  #     "DistToBoats" = "Boating",
  #     "DistToBottling" = "Bottling plants",
  #     "DistToColleges" = "Colleges",
  #     "DistToDistrib" = "Distribution centers",
  #     "DistToGarages" = "Garages",
  #     "DistToIntermodal" = "Intermodal platforms",
  #     "DistToMail" = "Mail carriers",
  #     "DistToMarket" = "Farmers market",
  #     "DistToMoving" = "Moving companies",
  #     "DistToPeople" = "Popular destinations",
  #     "DistToPort" = "Port",
  #     "DistToRail" =  "Railroad",
  #     "DistToRoad" = "Major road",
  #     "DistToTruckStop" = "Truck stops",
  #     "DistToWineries" = "Wineries",
  #     "DistToWood" = "Wood-related activities"))) +
  # # coord_cartesian(ylim=c(0, 30000)) +
  # xlab("Category") + ylab("Distance to the nearest... (km)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  


## FOR MATT SLF 101 MEDIANS FULL
random_long_median <- random_long %>% filter(grepl("_median", DistanceType))
head(random_long_median)
random_long_median$DistanceType <- as.factor(random_long_median$DistanceType)
levels(random_long_median$DistanceType)
levels(random_long_median$DistanceType) <- gsub(pattern = "_median", replacement = "", x = levels(random_long_median$DistanceType))


head(random_long_median)
summary_sim_medianfull <- random_long_median %>% group_by(Category_full, DistanceType) %>% summarise(median_sim = median(DistanceValue), sd_sim = sd(DistanceValue))
head(summary_sim_medianfull)



slf_obsmedians <- slf_cat_long %>% 
  group_by(Category_full, DistanceType) %>% 
  summarise(MedianDistance = median(DistanceValue)#,
            # MedianDistance = median(DistanceValue)
            )



summary_medianfull <- merge(summary_sim_medianfull, slf_obsmedians)
summary_medianfull %<>% rename(median_obs = MedianDistance) %>% 
  mutate(effect_size = (median_obs - median_sim)/sd_sim)
write.csv(summary_medianfull, "summary_simulations_median_full_transports.csv")


summary_medianfull_cat <- merge(summary_medianfull, Names)
write.csv(summary_medianfull_cat, "summary_simulations_median_full_transports.csv")
```

```{r slf 101 rarefied}

## FOR MATT SLF 101 MEAN RAREFIED
# Simulated means
MeanDistances <- read.csv(file.path(here(), "exported-data", "Distances_random distrib", "SimulatedMeans_rarefiedchull.csv"))
names(MeanDistances)
head(MeanDistances)

#Modify dataset to get distribution of distances with a column for the type of distance, and one column for the type of dataset (Full or Rarefied)
random_long <- MeanDistances %>%
  pivot_longer(cols = starts_with("DistTo"), names_to = "DistanceType", values_to = "DistanceValue")

random_long_mean <- random_long %>% filter(grepl("_mean", DistanceType))
head(random_long_mean)
random_long_mean$DistanceType <- as.factor(random_long_mean$DistanceType)
levels(random_long_mean$DistanceType)
levels(random_long_mean$DistanceType) <- gsub(pattern = "_mean", replacement = "", x = levels(random_long_mean$DistanceType))


head(random_long_mean)
summary_sim_rare <- random_long_mean %>% group_by(Category_rarefied, DistanceType) %>% summarise(mean_sim = mean(DistanceValue), sd_sim = sd(DistanceValue))
head(summary_sim_rare)



slf_obsmeans <- slf_cat_long %>% 
  group_by(Category_rare, DistanceType) %>% 
  summarise(MeanDistance = mean(DistanceValue)#,
            # MedianDistance = median(DistanceValue)
            ) %>% 
  rename(Category_rarefied = Category_rare)




summary_rare <- merge(summary_sim_rare, slf_obsmeans)
summary_rare %<>% rename(mean_obs = MeanDistance) %>% 
  mutate(effect_size = (mean_obs - mean_sim)/sd_sim)
write.csv(summary_rare, "summary_simulations_rare.csv")


summary_rare_cat <- merge(summary_rare, Names)
write.csv(summary_rare_cat, "summary_simulations_mean_rare_transports.csv")


## FOR MATT SLF 101 MEDIANS FULL
random_long_median <- random_long %>% filter(grepl("_median", DistanceType))
head(random_long_median)
random_long_median$DistanceType <- as.factor(random_long_median$DistanceType)
levels(random_long_median$DistanceType)
levels(random_long_median$DistanceType) <- gsub(pattern = "_median", replacement = "", x = levels(random_long_median$DistanceType))


head(random_long_median)
summary_sim_medianrare <- random_long_median %>% group_by(Category_rarefied, DistanceType) %>% summarise(median_sim = median(DistanceValue), sd_sim = sd(DistanceValue))
head(summary_sim_medianrare)



slf_obsmedians <- slf_cat_long %>% 
  group_by(Category_rare, DistanceType) %>% 
  summarise(MedianDistance = median(DistanceValue)#,
            # MedianDistance = median(DistanceValue)
            ) %>% 
  rename(Category_rarefied = Category_rare)



summary_medianrare <- merge(summary_sim_medianrare, slf_obsmedians)
summary_medianrare %<>% rename(median_obs = MedianDistance) %>% 
  mutate(effect_size = (median_obs - median_sim)/sd_sim)
write.csv(summary_medianrare, "summary_simulations_median_rare.csv")

summary_rare_mediancat <- merge(summary_medianrare, Names)
write.csv(summary_rare_mediancat, "summary_simulations_median_rare_transports.csv")
```


We can visualize the results on Figure 2. The histogram represents the distribution of distances under the null hypothesis of random dispersal of jumpers, for each type of transport. The black vertical lines indicate the significance limits. An observed value situated outside of these vertical lines leads to the rejection of the null hypothesis. The red line indicates the average distance to transports observed in our dataset.  
For all three types of transports, the observed location of jumpers is significantly closer to transports than random.  


\newpage

# 6. Results and conclusion

In this vignette, we discovered that:  

(0) The 40,988 surveys of the dataset that confirmed either the presence of an established SLF population or the absence of SLF have been conducted on average 13, 17 and 19 km away from major roads, railroads and airports, respectively. It is much further than random. *==> The survey method is not biased towards transport infrastructures*

(1) SLF populations, both from jump events and diffusive spread, are not located randomly, but very significantly close to transport infrastructures: roads, rail and airports. The difference to the random distribution is the highest for railroads, then roads, then airports.  *==> SLF presence is tightly linked to transport infrastructures*

(2) Jump events are situated even closer to transport infrastructures than the other SLF populations, on average 444, 659 and 6,566 m away from major roads, rails and airports, respectively. *==> the establishment of new satellite populations is even more linked to transport infrastructures than diffusive spread*

(3) On the other hand, locations where SLF are not found are situated further than random from transport infrastructures. *==> SLF are unlikely to establish far from transport infrastructures*


Given these results, it is very likely that there is a causal relationship between SLF presence and transport infrastructures. SLF are likely transported by vehicles or their content, either as egg masses laid onto random materials, or at another life stage that crawls or flies on vehicles.
It is likely that some transports are more involved in SLF dispersal than others. Correlation is not causation: the proximity to some transports might appear significantly related to SLF presence only because these transports are found in overall highly connected areas. This might be the case for airports, because the correlation between SLF and airports appears looser than with roads or rails. Airports are typically in areas well connected by roads and rails. More, it is assumed that adult SLF cannot survive a flight.


In the next vignette, we are going to apply this knowledge to the spatial analysis of the northeastern US, by looking at areas considered as high risk of invasion (their proximity to SLF populations and to transport infrastructures), and projecting areas likely to be invaded sooner or later.


