---
title: "Predictions of slf jumps - buffers around landscape features"
author: "Nadege Belouard"
date: "4/11/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

We will use the knowledge we have on the location of SLF jumps location so far to project potential areas of colonization.

# 1. Setup

## Load packages
```{r setup for rendering, include = F, messages = F, warning = F}

# attaching necessary packages
library(tidyverse)
library(magrittr)
library(sf)
library(maps)
library(DescTools)
library(reshape2)
library(ggplot2)
library(here)
library(slfjumps)
library(leaflet)
library(purrr)
library(geosphere)

sf::sf_use_s2(FALSE)
```

## States map
```{r states names and centroid for global map, message = FALSE, warning = FALSE, echo = params$display}
# extracts a map of the States and recodes state labels to show the two-letter code rather than the full state name.

# obtaining simple feature objects for states and finding centroids for label positioning
states <- sf::st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))
# states <- st_transform(states, crs = "ESRI:102010")
# sf::st_as_sf(maps::map("county", plot = TRUE, fill = FALSE))
states <- cbind(states, st_coordinates(st_centroid(states)))

# making table key for state 2-letter abbreviations
# the vectors state.abb and state.name contains strings of all
# US states and abbreviations
state_abbr <- tibble(state.name = str_to_lower(state.name), state.abb) %>%
  left_join(tibble(ID = states$ID), ., by = c(ID = "state.name")) %>%
  mutate(state.abb = replace_na(state.abb, ""))

# adding 2-letter codes to sf
states$code <- state_abbr$state.abb


ggplot() +
    geom_sf(data = states, alpha = 0) +
  geom_text(data = states, aes(X, Y, label = code), size = 2) 


US <- st_read(file.path(here(), "figures", "GIS", "gadm36_Cont_USA_county.shp"), crs = "EPSG:4326", quiet = T)
st_crs(US)
US <- st_transform(US, crs = "ESRI:102010")

```



## Load files
```{r load datasets for analysis}
slf_uptodate <- read.csv(file.path(here(), "exported-data", "slf_obs_uptodate.csv"), h=T)
slfPA_uptodate <- read.csv(file.path(here(), "exported-data", "slfPA_uptodate.csv"), h=T)
hull <- st_read(file.path(here(), "figures", "GIS", "chull.shp"), quiet = T)
st_area(hull) #173,517,523,874 m2 = 173,517 km2
Thresholds <- read.csv(file.path(here(), "exported-data", "thresholds.csv"))
```

Map points
```{r map points}

# take data on all diffusers
diffusers <- slf_uptodate %>% filter(Category_full == "Diffusers")

diffusers <- st_as_sf(x = diffusers, coords = c("longitude_rounded", "latitude_rounded"), crs = "EPSG:4269", remove = F)

# take data on all thresholds/invasion fronts
head(Thresholds)
thresholds <- st_as_sf(x = Thresholds %>% filter(bio_year == 2020), coords = c("longitude_rounded", "latitude_rounded"), crs = "EPSG:4269", remove = F) 

# map them
ggplot(data = states) +
  geom_sf() +
  geom_sf(data = diffusers, col = "red") +
  geom_sf(data = thresholds) +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)
```

# 2. Buffer around invasion front
We want to create a buffer around the invasion front that represents the predicted zone where jumps can happen, to subset the zone in the US that might be at risk.


```{r distance to threshold observed}
DistToThreshold <- read.csv(file.path(here(), "exported-data", "DistToThreshold.csv"))
head(DistToThreshold)

DistToThreshold %>% group_by(bio_year, Rarefied) %>% summarise(maxdist = max(DistToThreshold),                                                     dist95 = quantile(DistToThreshold, 0.95))

# Rarefied dataset, up to 2020
Full <- DistToThreshold %>% filter(Rarefied == "Full")
Rarefied <- DistToThreshold %>% filter(Rarefied == "Rarefied")
dim(Rarefied)
max(Rarefied$DistToThreshold) # 307.8 km
mean(Rarefied$DistToThreshold) # 87.56 km
quantile(Rarefied$DistToThreshold, 0.95) # 262.55 km
quantile(Full$DistToThreshold, 0.95) #222.5 km

# Rarefied dataset, up to 2019
Full <- DistToThreshold %>% filter(Rarefied == "Full", bio_year %in% c(2014:2019))
Rarefied <- DistToThreshold %>% filter(Rarefied == "Rarefied", bio_year %in% c(2014:2019))
dim(Rarefied)
max(Rarefied$DistToThreshold) # 307.84
mean(Rarefied$DistToThreshold) # 87.78 km
quantile(Rarefied$DistToThreshold, 0.95) # 285.43 km
quantile(Full$DistToThreshold, 0.95) # 222.90 km
```

The distance of 95% of the jumps is 263 km (rarefied dataset), and the average distance is 88 km (rarefied dataset). 
So we will use a buffer of 263 km for predictions up to 2020.
For predictions up to 2019, the 95th percentile is 285.43 km

```{r buffer around invasion front 2020}
# We first need to create a polyon that represents the invasion front. It's not a minimum convex hull!
# We do this step on ArcGIS because it is easier than on sf
invasion_front <- st_read(file.path(here(), "figures", "GIS", "invasion_front.shp"), quiet = T)
st_crs(invasion_front)

# map that
ggplot(data = states) +
  geom_sf() +
  geom_sf(data = invasion_front, fill = "white") +
  # geom_sf(data = diffusers, col = "red") +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)

# Create a buffer around it
invasion_front <- st_transform(invasion_front, crs = "ESRI:102010")
buffer_invasion_front <- st_buffer(invasion_front, dist = 263000)


ggplot(data = states) +
  geom_sf() +
  geom_sf(data = buffer_invasion_front) +
  # geom_sf(data = diffusers) +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)


# Remove the area that is already colonized
atrisk_zone <- st_difference(buffer_invasion_front, invasion_front)
st_crs(atrisk_zone)

# Map it
ggplot(data = states) +
  geom_sf() +
  geom_sf(data = atrisk_zone, fill = "blue", alpha = 0.5) +
  # geom_sf(data = diffusers) +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)


# Keep the zone that is on land only!
US <- st_read(file.path(here(), "figures", "GIS", "gadm36_Cont_USA_county.shp"), crs = "EPSG:4326", quiet = T)
st_crs(US)
US <- st_transform(US, crs = "ESRI:102010")
atrisk_zone <- st_intersection(atrisk_zone, US) %>% st_union()

st_area(atrisk_zone)

st_write(atrisk_zone, file.path(here(), "figures", "GIS", "atrisk_zone.shp"), append = F)
```

```{r buffer around invasion front 2019}
# We first need to create a polyon that represents the invasion front. It's not a minimum convex hull!
# We do this step on ArcGIS because it is easier than on sf
invasion_front2019 <- st_read(file.path(here(), "figures", "GIS", "invasionfront_2019.shp"), quiet = T)
st_crs(invasion_front2019)

# map that
ggplot(data = states) +
  geom_sf() +
  geom_sf(data = invasion_front2019, fill = "blue", alpha = 0.5) +
  # geom_sf(data = invasion_front, fill = "green", alpha= 0.5) +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)

# Create a buffer around it
invasion_front2019 <- st_transform(invasion_front2019, crs = "ESRI:102010")
buffer_invasion_front2019 <- st_buffer(invasion_front2019, dist = 285400)


ggplot(data = states) +
  geom_sf() +
  geom_sf(data = buffer_invasion_front2019) +
  # geom_sf(data = diffusers) +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)


# Remove the area that is already colonized
atrisk_zone2019 <- st_difference(buffer_invasion_front2019, invasion_front2019)

# Map it
ggplot(data = states) +
  geom_sf() +
  geom_sf(data = atrisk_zone2019, fill = "blue", alpha = 0.5) +
  geom_sf(data = atrisk_zone, fill = "green", alpha = 0.5) +
  coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)

# Keep the zone that is on land only!
US <- st_read(file.path(here(), "figures", "GIS", "gadm36_Cont_USA_county.shp"), crs = "EPSG:4326", quiet = T)
st_crs(US)
US <- st_transform(US, crs = "ESRI:102010")

ggplot(data = US) +
  geom_sf()

atrisk_zone2019 <- st_intersection(atrisk_zone2019, US) %>% st_union()

st_area(atrisk_zone2019)/1e6

st_write(atrisk_zone2019, file.path(here(), "figures", "GIS", "atrisk_zone2019.shp"), append = F)

st_area(atrisk_zone2019)/sum(st_area(US))
length(st_area(US))
```



# 3. Intersect landscape features and at risk zone

## Load all landscape features
And create buffer around them
```{r load landscape features}

# Mail carriers
mail <- st_read(file.path(here(), "figures", "GIS", "landscape", "mail_buffer.shp"), quiet = T)

# Wood activities
wood <- st_read(file.path(here(), "figures", "GIS", "landscape", "wood_buffer.shp"), quiet = T)

# Wineries
wineries <- st_read(file.path(here(), "figures", "GIS", "landscape", "wineries_buffer.shp"), quiet = T)

# People
people <- st_read(file.path(here(), "figures", "GIS", "landscape", "people_buffer.shp"), quiet = T)

# Garages
garages <- st_read(file.path(here(), "figures", "GIS", "landscape", "garages_buffer.shp"), quiet = T)

# Boats
boats <- st_read(file.path(here(), "figures", "GIS", "landscape", "boats_buffer.shp"), quiet = T)

# Airports
airports <- st_read(file.path(here(), "figures", "GIS", "airports_primary_buffer.shp"), quiet = T)
airports <- st_transform(airports, crs = "ESRI:102010")

# Rail
rail <- st_read(file.path(here(), "figures", "GIS", "Railways_full_buffer5m.shp"), quiet = T)

# Road
road <- st_read(file.path(here(), "figures", "GIS", "road_buffer15m.shp"), quiet = T)

# Intersect of rail and road
roadrail_intersect <- st_read(file.path(here(), "figures", "GIS", "roadrail100m_union.shp"), quiet = T)
roadrail_intersect <- st_transform(roadrail_intersect, crs = "ESRI:102010")

```


## Intersect all landscape layers with at risk zone
For 2020
```{r intersect with at risk zone}

# At risk zone
atrisk_zone <- st_read(file.path(here(), "figures", "GIS", "atrisk_zone.shp"), quiet = T)

mail_zone <- st_intersection(mail, atrisk_zone) %>% 
  st_union()
st_write(mail_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "mail_zone.shp"), quiet = T, append = F)

wood_zone <- st_intersection(wood, atrisk_zone) %>% 
  st_union()
st_write(wood_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "wood_zone.shp"), quiet = T, append = F)

wineries_zone <- st_intersection(wineries, atrisk_zone) %>% 
  st_union()
st_write(wineries_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "wineries_zone.shp"), quiet = T, append = F)

people_zone <- st_intersection(people, atrisk_zone) %>% 
  st_union()
st_write(people_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "people_zone.shp"), quiet = T, append = F)

garages_zone <- st_intersection(garages, atrisk_zone) %>% 
  st_union()
st_write(garages_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "garages_zone.shp"), quiet = T, append = F)

boats_zone <- st_intersection(boats, atrisk_zone) %>% 
  st_union()
st_write(boats_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "boats_zone.shp"), quiet = T, append = F)

airports_zone <- st_intersection(airports, atrisk_zone) %>% 
  st_union()
st_write(airports_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "airports_zone.shp"), quiet = T, append = F)

rail_zone <- st_intersection(rail, atrisk_zone) %>% 
  st_union()
st_write(rail_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "rail_zone.shp"), quiet = T, append = F)

road_zone <- st_intersection(road, atrisk_zone) %>% 
  st_union()
st_write(road_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "road_zone.shp"), quiet = T, append = F)

roadrail_intersect_zone <- st_intersection(roadrail_intersect, atrisk_zone) %>% 
  st_union()
st_write(roadrail_intersect_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "roadrail_intersect_zone.shp"), quiet = T, append = F)


# Map it
ggplot(data = states) +
  geom_sf() +
  geom_sf(data = atrisk_zone, fill = "blue", alpha = 0.5) #+
  # geom_sf(data = roadrail_intersect_zone) +
  # coord_sf(xlim = c(1000000, 2000000), ylim = c(-200000, 600000), expand = FALSE)

```

For 2019 (might need to rerun the ones that are not in the manuscript, there was a mishap with atrisk_zone and atrisk_zone2019)
```{r intersect with at risk zone 2019}

# At risk zone
atrisk_zone2019 <- st_read(file.path(here(), "figures", "GIS", "atrisk_zone2019.shp"), quiet = T)

mail_zone <- st_intersection(mail, atrisk_zone2019) %>% 
  st_union()
st_write(mail_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "mail_zone2019.shp"), quiet = T, append = F)

wood_zone <- st_intersection(wood, atrisk_zone2019) %>% 
  st_union()
st_write(wood_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "wood_zone2019.shp"), quiet = T, append = F)

wineries_zone <- st_intersection(wineries, atrisk_zone2019) %>% 
  st_union()
st_write(wineries_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "wineries_zone2019.shp"), quiet = T, append = F)

people_zone <- st_intersection(people, atrisk_zone2019) %>% 
  st_union()
st_write(people_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "people_zone2019.shp"), quiet = T, append = F)

garages_zone <- st_intersection(garages, atrisk_zone2019) %>% 
  st_union()
st_write(garages_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "garages_zone2019.shp"), quiet = T, append = F)

boats_zone <- st_intersection(boats, atrisk_zone2019) %>% 
  st_union()
st_write(boats_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "boats_zone2019.shp"), quiet = T, append = F)

airports_zone <- st_intersection(airports, atrisk_zone2019) %>% 
  st_union()
st_write(airports_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "airports_zone2019.shp"), quiet = T, append = F)

rail_zone <- st_intersection(rail, atrisk_zone2019) %>% 
  st_union()
st_write(rail_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "rail_zone2019.shp"), quiet = T, append = F)

road_zone <- st_intersection(road, atrisk_zone2019) %>% 
  st_union()
st_write(road_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "road_zone2019.shp"), quiet = T, append = F)

roadrail_intersect_zone <- st_intersection(roadrail_intersect, atrisk_zone2019) %>% 
  st_union()
st_write(roadrail_intersect_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "roadrail_intersect_zone2019.shp"), quiet = T, append = F)

```


# 4. Create maps!

## Load packages
```{r setup for rendering, include = F, messages = F, warning = F}

# attaching necessary packages
library(tidyverse)
library(magrittr)
library(sf)
library(DescTools)
library(here)
library(leaflet)
library(purrr)
library(geosphere)
```

## Load data
For 2020
```{r load landscape features}
# Jump data
slf_uptodate <- read.csv(file.path(here(), "exported-data", "slf_obs_uptodate.csv"), h=T)
slfPA_uptodate <- read.csv(file.path(here(), "exported-data", "slfPA_uptodate.csv"), h=T)

# At risk zone
atrisk_zone <- st_read(file.path(here(), "figures", "GIS", "atrisk_zone.shp"), quiet = T)

mail_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "mail_zone.shp"), quiet = T)

wood_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "wood_zone.shp"), quiet = T)

wineries_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "wineries_zone.shp"), quiet = T)

people_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "people_zone.shp"), quiet = T)

garages_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "garages_zone.shp"), quiet = T)

boats_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "boats_zone.shp"), quiet = T)

airports_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "airports_zone.shp"), quiet = T)

rail_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "rail_zone_simple.shp"), quiet = T)
# rail_zone <- st_simplify(rail_zone)
# st_write(rail_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "rail_zone_simple.shp"))

road_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "road_zone.shp"), quiet = T)

roadrail_intersect_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "roadrail_intersect_zone.shp"), quiet = T)
```

For 2019
```{r load landscape features}
# Jump data
slf_uptodate <- read.csv(file.path(here(), "exported-data", "slf_obs_uptodate.csv"), h=T)
slfPA_uptodate <- read.csv(file.path(here(), "exported-data", "slfPA_uptodate.csv"), h=T)

# At risk zone
atrisk_zone <- st_read(file.path(here(), "figures", "GIS", "atrisk_zone2019.shp"), quiet = T)
st_area(atrisk_zone)

mail_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "mail_zone2019.shp"), quiet = T)

wood_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "wood_zone2019.shp"), quiet = T)

wineries_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "wineries_zone2019.shp"), quiet = T)

people_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "people_zone2019.shp"), quiet = T)

garages_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "garages_zone2019.shp"), quiet = T)

boats_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "boats_zone2019.shp"), quiet = T)

airports_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "airports_zone2019.shp"), quiet = T)

rail_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "rail_zone2019.shp"), quiet = T)
# rail_zone <- st_simplify(rail_zone)
# st_write(rail_zone, file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "rail_zone_simple2019.shp"))

road_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "road_zone2019.shp"), quiet = T)

roadrail_intersect_zone <- st_read(file.path(here(), "figures", "GIS", "landscape_zoneatrisk", "roadrail_intersect_zone2019.shp"), quiet = T)
```

Steps: 
(0) choose layer and risk level
(1) calculate distance with risk level
(2) make buffer with this distance

```{r make landscape layer buffers}

#load the distances
jumps <- read.csv(file.path(here(), "exported-data", "jumps.csv"))
dim(jumps)
jumps_dist <- slf_uptodate %>% filter(Category_full == "Jumpers")
dim(jumps_dist)
past_jumps <- jumps %>% left_join(jumps_dist)
dim(past_jumps)
jumpsPA_dist <- slfPA_uptodate %>% filter(Category_full == "Jumpers")
past_jumps <- past_jumps %>% left_join(jumpsPA_dist)
dim(past_jumps)

past_jumps %<>% filter(bio_year < 2020)

# jumps <- slf_uptodate %>% filter(Category_rare == "Jumpers")
# jumpsPA <- slfPA_uptodate %>% filter(Category_rare == "Jumpers")
# jumps <- merge(jumps, jumpsPA, all = T)
# write.csv(jumps, file.path(here(), "..", "slfjump_app", "data", "jumps.csv"))

dist <- past_jumps[which(names(past_jumps) == 'DistToRoad')] #choose layer
risk = 0.95 #choose risk level
threshold <- quantile(dist[1], probs = risk, na.rm = T) #calculate distance

buffer <- st_buffer(road_zone, dist = threshold) #choose layer and make buffer
st_area(buffer)/sum(st_area(US))


# #for 2021
# dist <- jumps_dist[which(names(jumps_dist) == 'DistToRoad')]
# risk = 0.95
# threshold <- quantile(dist[1], probs = risk, na.rm = T)
# 
# buffer <- st_buffer(road_zone, dist = threshold)
# st_area(buffer)/1e6
# /sum(st_area(US))
# 185024000000/sum(st_area(US)) #rail
# 202499000000/sum(st_area(US)) #road
# 128462000000/sum(st_area(US)) #intersect
# 
# st_area(atrisk_zone)/sum(st_area(US))

atrisk_zone <- atrisk_zone %>% st_transform(crs = "+proj=longlat +datum=WGS84") 
buffer <- buffer %>% st_transform(crs = "+proj=longlat +datum=WGS84") 

```

Map it! With leaflet: requires to transform layers to WGS84
```{r jumps map}

leaflet(data = buffer) %>% 
  addTiles() %>% 
  addPolygons() %>% 
  # addPolygons(data = rail_zone, col = "black") %>%
  addPolygons(data = atrisk_zone, fillColor = "red", weight = 0.1)

```

Now add jumps realized in 2020 and see if they are in the predicted high risk zone!
```{r jumps map}

jumps2020 <- jumps %>% filter(bio_year == 2020)
jumps2020 <- st_as_sf(x = jumps2020, coords = c("longitude_rounded", "latitude_rounded"), crs = "EPSG:4269", remove = F) %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") 
print(jumps2020)

buffer %<>% 
  st_transform(crs = "+proj=longlat +datum=WGS84")

leaflet(data = buffer) %>% 
  addTiles() %>% 
  addPolygons() %>%
  addCircleMarkers(data = jumps2020, ~longitude_rounded, ~latitude_rounded, 
             color = "red", fillColor = "red", fillOpacity = 0.7, radius = 7
             ) #%>%
  addPolygons(data = atrisk_zone, fillColor = "red", weight = 0.1)

```


With ggplot2 (for manuscript)
```{r map predictions for 2020}

jumps2020 <- jumps %>% filter(bio_year == 2020)
jumps2020 <- st_as_sf(x = jumps2020, coords = c("longitude_rounded", "latitude_rounded"), crs = "EPSG:4269", remove = F)
#%>%
 # st_transform(crs = "+proj=longlat +datum=WGS84")

# Roads
dist <- past_jumps[which(names(past_jumps) == 'DistToRoad')] #choose layer
risk = 0.95 #choose risk level
threshold <- quantile(dist[1], probs = risk, na.rm = T) #calculate distance

buffer <- st_buffer(road_zone, dist = threshold) #choose layer and make buffer
st_area(buffer) #202,500
st_area(buffer)/sum(st_area(US)) #0.0270


map_predictionsroad <- ggplot() +
  geom_sf(data = states, fill = "white") +
  geom_sf(data = atrisk_zone, fill = "yellow", col = "yellow", alpha = 0.3,) +
  geom_sf(data = buffer, fill = "red", col = "red", alpha = 0.3) +
  geom_sf(data = jumps2020, size = 3) +
    geom_sf(data = states, alpha = 0) +
  coord_sf(xlim = c(-83, -70), ylim = c(36, 45)) +
  
  labs(x = "Longitude", y = "Latitude") +
  ggtitle("(A) Major roads")

map_predictionsroad
ggsave(file.path(here(), "figures", "manuscript figures", "S4.buffer roads.jpg"), map_predictionsroad, height = 8, width = 8)


# Railroads
dist <- past_jumps[which(names(past_jumps) == 'DistToRail')] #choose layer
risk = 0.95 #choose risk level
threshold <- quantile(dist[1], probs = risk, na.rm = T) #calculate distance

buffer <- st_buffer(rail_zone, dist = threshold) #choose layer and make buffer
st_area(buffer) #174,34
st_area(buffer)/sum(st_area(US))

map_predictionsrail <- ggplot() +
  geom_sf(data = states, fill = "white") +
  geom_sf(data = atrisk_zone, fill = "yellow", col = "yellow", alpha = 0.3,) +
  geom_sf(data = buffer, fill = "red", col = "red", alpha = 0.3) +
  geom_sf(data = jumps2020, size = 3) +
    geom_sf(data = states, alpha = 0) +
  coord_sf(xlim = c(-83, -70), ylim = c(36, 45)) +
  
  labs(x = "Longitude", y = "Latitude") +
  ggtitle("(B) Railroads")

map_predictionsrail
ggsave(file.path(here(), "figures", "manuscript figures", "S4.buffer rail.jpg"), map_predictionsrail, height = 8, width = 8)


# Intersection rail/road
dist <- past_jumps[which(names(past_jumps) == 'DistIntRlRd')] #choose layer
risk = 0.95 #choose risk level
threshold <- quantile(dist[1], probs = risk, na.rm = T) #calculate distance

buffer <- st_buffer(roadrail_intersect_zone, dist = threshold) #choose layer and make buffer
st_area(buffer) # 121,116 km2
st_area(buffer)/sum(st_area(US)) #1.6%

map_predictionsint <- ggplot() +
  geom_sf(data = states, fill = "white") +
  geom_sf(data = atrisk_zone, fill = "yellow", col = "yellow", alpha = 0.3,) +
  geom_sf(data = buffer, fill = "red", col = "red", alpha = 0.3) +
  geom_sf(data = jumps2020, size = 3) +
    geom_sf(data = states, alpha = 0) +
  coord_sf(xlim = c(-83, -70), ylim = c(36, 45)) +
  
  labs(x = "Longitude", y = "Latitude") +
  ggtitle("(C) Intersection of roads and railroads")

map_predictionsint
ggsave(file.path(here(), "figures", "manuscript figures", "S4.buffer intersect.jpg"), map_predictionsint, height = 8, width = 8)



# Package carriers
dist <- past_jumps[which(names(past_jumps) == 'DistToMail')] #choose layer
risk = 0.95 #choose risk level
threshold <- quantile(dist[1], probs = risk, na.rm = T) #calculate distance

buffer <- st_buffer(mail_zone, dist = threshold) #choose layer and make buffer
st_area(buffer) # 52,729 km2
st_area(buffer)/sum(st_area(Pennsylvania)) #0.7%

Pennsylvania <- US %>% filter(NAME_1 == "Pennsylvania") %>% st_union()
atrisk_PA <- st_intersection(atrisk_zone, Pennsylvania)
buffer_PA <- st_intersection(buffer, Pennsylvania)
jumps2020 %<>% st_transform(crs = st_crs(Pennsylvania))
jumps_PA <- st_intersection(jumps2020, Pennsylvania)


map_predictionsmail <- ggplot() +
  geom_sf(data = states, fill = "white") +
  geom_sf(data = atrisk_PA, fill = "yellow", col = "yellow", alpha = 0.3,) +
  geom_sf(data = buffer_PA, fill = "red", col = "red", alpha = 0.3) +
  geom_sf(data = jumps_PA, size = 3) +
    geom_sf(data = states, alpha = 0) +
  coord_sf(xlim = c(-83, -70), ylim = c(36, 45)) +
  
  labs(x = "Longitude", y = "Latitude") +
  ggtitle("(D) Package carriers in Pennsylvania")

map_predictionsmail
ggsave(file.path(here(), "figures", "manuscript figures", "S4.buffer mail.jpg"), map_predictionsmail, height = 8, width = 8)


figure4 <- grid.arrange(map_predictionsroad,
                        map_predictionsrail,
             map_predictionsint,
             map_predictionsmail)
ggsave(file.path(here(), "figures", "manuscript figures", "S4.buffer figure.jpg"), figure4, height = 8, width = 8)
```